---
title: "R Notebook"
output: html_notebook
---

#### Packages
```{r}
here::i_am("Scripts/Rmd_Scripts/4_Occupancy_analyses.Rmd")

library(tidyverse)
library(zetadiv)
library(tidyverse)
library(ggpubr)
library(mgcv)
library(mgcViz)
library(grid)
library(gridExtra)
library(here)
library(knitr)
library(conflicted)
library(tictoc)

conflicts_prefer(dplyr::select)
conflicts_prefer(knitrProgressBar::progress_estimated)
```

#### Functions
```{r}
#I have tweaked the Zeta.order.ex function to include the varmat in the output

Zeta.order.ex.2 <- function (data.spec, order = 1, sd.correct = TRUE, rescale = FALSE, 
    empty.row = "empty") 
{
    if (empty.row == "remove") {
        data.spec <- data.spec[-which(rowSums(data.spec) == 0), 
            ]
    }
    if (order > dim(data.spec)[1]) {
        stop("Error: wrong value for \"order\": it must be equal or lower than the number of sites.")
    }
    data.spec <- as.matrix(data.spec)
    intercept_mat <- t(data.spec) %*% data.spec
    occupancy <- colSums(data.spec)
    p <- exp(lchoose(occupancy, order) - lchoose(nrow(data.spec), 
        order))
    zeta.val <- sum(p)
    varmat <- exp(lchoose(intercept_mat, order) - lchoose(nrow(data.spec), 
        order))
    for (j in 1:length(occupancy)) {
        for (k in 1:length(occupancy)) {
            varmat[j, k] <- varmat[j, k] - p[j] * p[k]
        }
    }
    if (sd.correct == TRUE) {
        zeta.val.sd <- sqrt(sum(varmat) * choose(nrow(data.spec), 
            order)/(choose(nrow(data.spec), order) - 1))
    }
    else {
        zeta.val.sd <- sqrt(sum(varmat))
    }
    zeta.order <- list()
    zeta.order$zeta.order <- order
    zeta.order$combinations <- choose(x <- dim(data.spec)[1], 
        order)
    zeta.order$zeta.val <- zeta.val
    zeta.order$zeta.val.sd <- zeta.val.sd
    zeta.order$varmat <- varmat
        return(zeta.order)
}


# a function to get the rowsums for each species in the varmat normalised by variance (sumo of whole matrix)and add to a dataframe which should already have the occupancy and possibly rowsums for other orders of zeta.
#normalised rowsums of the varmat are supposed to represent contribution of each species to variation in zeta diversity so that occupancy can be plotted against contribution to zeta for differnt orders of zeta

# arguments are: site by species matrix, zeta order, name for new variable of normalised rowsums of varmat for each species, data frame to merge the new variable into

get.var.share <- function(data.spec, ord, new.var.name, data.occ){
  
  z <-  Zeta.order.ex.2(data.spec, order = ord)


df1 <- data.frame(z$varmat)

df2<- data.frame(rowSums(df1)/sum(df1))

names(df2)<- new.var.name

out.df <- merge(data.occ, df2, by = "row.names")

out.df<- column_to_rownames(out.df, "Row.names")


return(out.df)

}

#occ.df1<- column_to_rownames(occ.df1, "Row.names")
```

#### File Paths
```{r}
data_path <- paste0("./Data/Processed_Data/Spatial/all_species.rds")

out_path <-  paste0(here("Output/Spatial/zeta_declines_decays/all_species/"))

if(!dir.exists(out_path)) dir.create(out_path, recursive = TRUE)
```

#### Load data
```{r}
data <- readRDS(here(data_path))

site.by.species <- data$site.by.species
  
site.by.env <- data$site.by.env

site.by.xy <- data$site.by.xy

species <- data$species
```

#### Recode Wider countryside sp to Habitat generalist
```{r}
species <- species %>% 
  mutate(
    strategy = case_match(strategy, "Wider countryside sp" ~ "Habitat generalist", .default = "Habitat specialist"))

```

#### Inspect occupancy
```{r}
# here I am trying to work out relative influence of different species on zeta diversity in relation to their occupancy

# "occ" is number of sites occupied
# "p.occ" is proportion of sites occupied

occ.df <- data.frame(cbind(colSums(site.by.species), colSums(site.by.species)/nrow(site.by.species)))
names(occ.df) <- c("occ", "p.occ")


occ.df <- occ.df %>% 
  mutate(rare.cat = case_when(p.occ < 0.2 ~ "1.rare",
                              p.occ >=0.2 & p.occ < 0.4 ~ "2.med.rare",
                              p.occ >=0.4 & p.occ < 0.6 ~ "3.med", 
                              p.occ >= 0.6 & p.occ < 0.8 ~ "4.med.common",
                              p.occ >= 0.8 ~ "5.common"))

hist(occ.df$p.occ, breaks = 20)

```

# Calculate expected contribution to zeta diversity of different butterfly species for different zeta orders, assuming random occurrence of all species
```{r}
# To anyone reading this code, I apologize for the prodigious use of copy and paste!

occ.df<- data.frame( cbind(colSums(site.by.species), colSums(site.by.species)/nrow(site.by.species))) %>%
  rownames_to_column(var = "species.no")

names(occ.df) <- c("species.no", "occ", "p.occ")

occ.df <- merge(species, occ.df, by = "species.no")


# here I am trying to work out relative influence of different species on zeta diversity in relation to their occupancy
 
# ex.share.n 
# First we calculate the probability that a species is shared between n sites in one sample but not shared between n sites in a second sample
# Then we normalise by column sums to obtain expected share of variance in zeta diversity contributed by each species
occ.df <- occ.df %>% 
  mutate(rare.cat = case_when(p.occ < 0.2 ~ "1.rare",
                              p.occ >=0.2 & p.occ < 0.4 ~ "2.med.rare",
                              p.occ >=0.4 & p.occ < 0.6 ~ "3.med", 
                              p.occ >= 0.6 & p.occ < 0.8 ~ "4.med.common",
                              p.occ >= 0.8 ~ "5.common"),
  # # binomial variance proportional to p(1-p) and here p is the probability of being in i sites
           ex.share.2 = (p.occ^2)*(1-p.occ^2),
         ex.share.3 = (p.occ^3)*(1-p.occ^3),
         ex.share.4 = (p.occ^4)*(1-p.occ^4),
         ex.share.5 = (p.occ^5)*(1-p.occ^5),
         ex.share.6 = (p.occ^6)*(1-p.occ^6),
         ex.share.7 = (p.occ^7)*(1-p.occ^7),
         ex.share.8 = (p.occ^8)*(1-p.occ^8), 
         ex.share.9 = (p.occ^9)*(1-p.occ^9),
         ex.share.10 = (p.occ^10)*(1-p.occ^10),
         ex.share.15 = (p.occ^15)*(1-p.occ^15),
         ex.share.20 = (p.occ^20)*(1-p.occ^20),
         ex.share.30 = (p.occ^30)*(1-p.occ^30),
        ex.share.40 = (p.occ^40)*(1-p.occ^40),
         ex.share.50 = (p.occ^50)*(1-p.occ^50)
         
    )

# normalise expected share of variance by column sums

occ.df[, 10:23] <- lapply(occ.df[, 10:23], function(x) x/sum(x))

# I need species no as row names again for functions below
occ.df<- column_to_rownames(occ.df, var = "species.no")
```


#### Plot expected contribution to zeta diversity of different butterfly species
```{r}

ggplot(occ.df, aes(x=p.occ, y=ex.share.2, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 2", x = "Occupancy", y = "Expected Variance Contribution")


ggplot(occ.df, aes(x=p.occ, y=ex.share.3, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 3", x = "Occupancy", y = "Expected Variance Contribution")


ggplot(occ.df, aes(x=p.occ, y=ex.share.4, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 4", x = "Occupancy", y = "Expected Variance Contribution")


ggplot(occ.df, aes(x=p.occ, y=ex.share.5, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 5", x = "Occupancy", y = "Expected Variance Contribution")



ggplot(occ.df, aes(x=p.occ, y=ex.share.6, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 6", x = "Occupancy", y = "Expected Variance Contribution")


ggplot(occ.df, aes(x=p.occ, y=ex.share.7, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 7", x = "Occupancy", y = "Expected Variance Contribution")



ggplot(occ.df, aes(x=p.occ, y=ex.share.8, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 8", x = "Occupancy", y = "Expected Variance Contribution")


ggplot(occ.df, aes(x=p.occ, y=ex.share.9, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 9", x = "Occupancy", y = "Expected Variance Contribution")


ggplot(occ.df, aes(x=p.occ, y=ex.share.10, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 10", x = "Occupancy", y = "Expected Variance Contribution")


ggplot(occ.df, aes(x=p.occ, y=ex.share.15, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 15", x = "Occupancy", y = "Expected Variance Contribution")


ggplot(occ.df, aes(x=p.occ, y=ex.share.20, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 20", x = "Occupancy", y = "Expected Variance Contribution")



ggplot(occ.df, aes(x=p.occ, y=ex.share.30, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 30", x = "Occupancy", y = "Expected Variance Contribution")


ggplot(occ.df, aes(x=p.occ, y=ex.share.40, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 40", x = "Occupancy", y = "Expected Variance Contribution")


ggplot(occ.df, aes(x=p.occ, y=ex.share.50, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 50", x = "Occupancy", y = "Expected Variance Contribution")
```

#### Extract observed share of variance of zeta diversity contributed by different species of butterfly for different zeta orders
```{r}
occ.df <- get.var.share(site.by.species, 2, "share.2", occ.df)
occ.df <- get.var.share(site.by.species, 3, "share.3", occ.df)
occ.df <- get.var.share(site.by.species, 4, "share.4", occ.df)
occ.df <- get.var.share(site.by.species, 5, "share.5", occ.df)
occ.df <- get.var.share(site.by.species, 6, "share.6", occ.df)
occ.df <- get.var.share(site.by.species, 7, "share.7", occ.df)
occ.df <- get.var.share(site.by.species, 8, "share.8", occ.df)
occ.df <- get.var.share(site.by.species, 9, "share.9", occ.df)
occ.df <- get.var.share(site.by.species, 10, "share.10", occ.df)
occ.df <- get.var.share(site.by.species, 15, "share.15", occ.df)
occ.df <- get.var.share(site.by.species, 20, "share.20", occ.df)
occ.df <- get.var.share(site.by.species, 30, "share.30", occ.df)
occ.df <- get.var.share(site.by.species, 40, "share.40", occ.df)
occ.df <- get.var.share(site.by.species, 50, "share.50", occ.df)
```

#### Plot share of variance of zeta diversity contributed by different species of butterfly
```{r}
ggplot(occ.df, aes(x=p.occ, y=share.2, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 2", x = "Occupancy", y = "Variance Contribution")


ggplot(occ.df, aes(x=p.occ, y=share.3, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 3", x = "Occupancy", y = "Variance Contribution")


ggplot(occ.df, aes(x=p.occ, y=share.4, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 4", x = "Occupancy", y = "Variance Contribution")


ggplot(occ.df, aes(x=p.occ, y=share.5, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 5", x = "Occupancy", y = "Variance Contribution")



ggplot(occ.df, aes(x=p.occ, y=share.6, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 6", x = "Occupancy", y = "Variance Contribution")


ggplot(occ.df, aes(x=p.occ, y=share.7, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 7", x = "Occupancy", y = "Variance Contribution")



ggplot(occ.df, aes(x=p.occ, y=share.8, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 8", x = "Occupancy", y = "Variance Contribution")


ggplot(occ.df, aes(x=p.occ, y=share.9, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 9", x = "Occupancy", y = "Variance Contribution")


ggplot(occ.df, aes(x=p.occ, y=share.10, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 10", x = "Occupancy", y = "Variance Contribution")


ggplot(occ.df, aes(x=p.occ, y=share.15, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 15", x = "Occupancy", y = "Variance Contribution")


ggplot(occ.df, aes(x=p.occ, y=share.20, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 20", x = "Occupancy", y = "Variance Contribution")



ggplot(occ.df, aes(x=p.occ, y=share.30, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 30", x = "Occupancy", y = "Variance Contribution")


ggplot(occ.df, aes(x=p.occ, y=share.40, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 40", x = "Occupancy", y = "Variance Contribution")


ggplot(occ.df, aes(x=p.occ, y=share.50, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 50", x = "Occupancy", y = "Variance Contribution")

```


#### Create individual panels for multipanel plot (selected from above two sets of plots)
```{r}
p1 <- ggplot(occ.df, aes(x=p.occ, y=ex.share.2, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 2", x = " ", y = "Expected Variance \nContribution") + 
  theme(legend.position="none")


p2 <- ggplot(occ.df, aes(x=p.occ, y=ex.share.4, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 4", x = " ", y = " ") + 
  theme(legend.position="none")

p3 <- ggplot(occ.df, aes(x=p.occ, y=ex.share.20, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title="Zeta = 20", x = " ", y = " ") + 
  theme(legend.position="none")



p4 <- ggplot(occ.df, aes(x=p.occ, y=share.2, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title=" ", x = "Occupancy", y = "Observed Variance \nContribution") + 
  theme(legend.position="none")


p5 <-ggplot(occ.df, aes(x=p.occ, y=share.4, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title=" ", x = "Occupancy", y = " ") + 
  theme(legend.position="none")

p6 <- ggplot(occ.df, aes(x=p.occ, y=share.20, fill=strategy)) + geom_point(shape=21, size = 3) + 
  scale_fill_manual(values=c("blue", "green")) +
   labs(title=" ", x = "Occupancy", y = " ") + 
  theme(legend.position="none")

plot_list <- list(p1, p2, p3, p4, p5, p6)
```

#### Make selected plots into a multipanel plot and write as png file
```{r}
nr <- 2
nc <- 3

top_heads <- c(" ", " ", " ")

# to include column caption as a separate row, I addapted code from https://stackoverflow.com/questions/45473843/put-row-and-column-titles-using-grid-arrange-in-r
all_elements <- rbind(tableGrob(t(top_heads), theme = ttheme_minimal(), rows = ""), 
                 cbind(tableGrob(rep(" ",nr), theme = ttheme_minimal()), # I couldn't work out how not to have row headings so made them a space 
                       arrangeGrob(grobs = plot_list, ncol = nc),  size = "last"), size = "last")
grid.newpage()
grid.draw(all_elements)


png(file = paste0(out_path, "/contribution_zeta_variance.png"), pointsize=10, width=1920, height=2800, res=300)
grid.newpage()
grid.draw(all_elements)
dev.off()
```


#### Plot rank occupancy
```{r}

occ.df$rank <- rank(-occ.df$p.occ, ties.method = "random")

hist(occ.df$p.occ, breaks = 20)

plot(occ.df$p.occ ~ occ.df$rank)
```

#### Output occupancy table and rank occupancy plot with different colours of habitat specialist and generalists
```{r}
occ.out <- occ.df %>%
  select(common.name, sci.name, strategy, habitat, occ, p.occ, rank) %>%
  arrange(desc(occ)) %>%
  rename("Common name" = common.name,
         "Scientific name" = sci.name,
         "Ecological category " = strategy,
         "Habitat" = habitat, 
         "Number of sites" = occ, 
         "Proportion of sites " = p.occ,
         "Rank occupancy" = rank)
 write_csv(occ.out, file = paste0(out_path, "/butterfly_occupancy.csv"))
kable(occ.out, digits = 3)

# I need a factor to use groups in ggplot2
occ.df <- occ.df %>% 
  mutate(strategy = factor(strategy))

# Make the Plot
rank_occ <- ggplot(occ.df, aes(x = rank, y = p.occ)) + 
         geom_point(color='black', shape=21, size=2, aes(fill = strategy)) +
        scale_fill_manual(values = c("blue", "green"),
                          name = "Butterfly\ngroup") +  
        ylab("Proportion of sites occupied") +
        xlab("Rank")
rank_occ

png(file = paste0(out_path, "/rank_occupancy.png"))
rank_occ
dev.off()
```