---
title: "R Notebook"
output: html_notebook
---

#### Setup and Packages
```{r}
#All the packes are used, checked24/04/2024

here::i_am("Scripts/Rmd_Scripts/butterflies_combine_env_data.Rmd")

library(tidyverse)
library(conflicted)
library(tictoc)
library(terra)
library(tidyterra)
library(sf)
library(tmap)
library(gridExtra)
library(ggpubr)
library(GGally)
library(here)

conflicts_prefer(dplyr::filter)
```
To do: check we arereading correct processed input here
       get rid of northern Ireland and Ireland (mask)
       #mask isle of man and channel islands

#### Load data
```{r}
# load polygon to use to crop and mask as map extents a bit different

gb <- readRDS(here("./Data/Maps/gb_multipolygon_simplified.rds"))

str(gb, max.level = 1)

crs(gb)

butterfly_data <- readRDS(here("./Data/Processed_Data/Spatial/spatial_ukbms_obs_sites_spp.rds"))

site.by.species <- butterfly_data$site.by.species

sites <- butterfly_data$sites

species <-  butterfly_data$species

site.by.humans <- readRDS(here("./Data/Masked_Cropped_Environmental_Data/site_by_humans.rds"))

site.by.ph <- readRDS(here("./Data/Masked_Cropped_Environmental_Data/site_by_ph.rds"))

site.by.trees <-  readRDS(here("./Data/Input_Environmental_Data/trees_1km_df.rds"))

site.by.nitrogen <- readRDS(here("./Data/Masked_Cropped_Environmental_Data/site_by_nitrogen.rds"))

site.by.pesticide.risk <- readRDS(here("./Data/Masked_Cropped_Environmental_Data/site_by_pesticide_risk.rds"))

site.by.wetness <- readRDS(here("./Data/Masked_Cropped_Environmental_Data/site_by_wetness.rds"))

site.by.temp <- readRDS(here("./Data/Masked_Cropped_Environmental_Data/site_by_temp.rds"))

site.by.prec <- readRDS(here("./Data/Masked_Cropped_Environmental_Data/site_by_prec.rds"))
```

#### Functions
```{r}
log10_plusone <- function(x)
{log10(x + 1)}

source(here("Scripts/R_scripts/butterfly_richness_turnover_functions.R"))
```

#### Rename Environmental variables and log transform where necessary
```{r}
tic()
site.by.temp <- site.by.temp %>%
  rename(mean.ann.temp = mean)%>%
    filter(is.na(mean.ann.temp)==F)

site.by.prec <- site.by.prec %>%
  rename(mean.ann.rain = mean) %>%
    filter(is.na(mean.ann.rain)==F)

site.by.ph <- site.by.ph %>%
  rename(ph = ph_idw)%>%
    filter(is.na(ph)==F)

site.by.humans <- site.by.humans %>%
  rename(humans = gbr_pd_2017_1km)%>%
    filter(is.na(humans)==F)

site.by.trees <- site.by.trees %>%
  rename(tree.density = tree.cover.density)%>%
    filter(is.na(tree.density)==F)

site.by.nitrogen <-select(site.by.nitrogen, x, y, total.n, ngr.1km)%>%
    filter(is.na(total.n)==F)

site.by.pesticide.risk <- site.by.pesticide.risk %>%
  rename(pesticide.risk = pest)%>%
    filter(is.na(pesticide.risk)==F)

site.by.wetness <- site.by.wetness %>%
  rename(topographic.wetness = Band1) %>%
    filter(is.na(topographic.wetness)==F)

# log base 10 transformations

site.by.log10.pesticide.risk <- site.by.pesticide.risk %>%
  mutate(log10.pesticide.risk = log10_plusone(pesticide.risk)) %>%
  select(x, y, log10.pesticide.risk, ngr.1km) #also gets it in correct order

site.by.log10.humans <- site.by.humans %>%
  mutate(log10.humans = log10_plusone(humans)) %>%
  select(x, y, log10.humans, ngr.1km) #also gets it in correct order

site.by.log10.mean.ann.rain <- site.by.prec %>%
  mutate(log10.mean.ann.rain = log10_plusone(mean.ann.rain)) %>%
  select(x, y, log10.mean.ann.rain, ngr.1km) #also gets it in correct order

```




#### Merge env variables
```{r}
# Make a list to be able to use lapply
layers <- list(site.by.prec, site.by.log10.mean.ann.rain, site.by.wetness, 
               site.by.ph, site.by.nitrogen, site.by.trees, site.by.pesticide.risk, site.by.log10.pesticide.risk, 
               site.by.humans, site.by.log10.humans)

# Remove x and y from 7 of the layers as this is not needed for merging
layers2 <- lapply(layers, select, -c(x, y)) 

# Make a datframe for merged data for plotting, including the first env var
comb <- merge(site.by.temp, layers2[[1]], by = "ngr.1km")

# Merge each remaining env var dataframe into the combined one
for (i in 2:length(layers2)){
  
  comb <- merge(comb, layers2[[i]], by = "ngr.1km")
}


```

```{r}
comb.check <- select(comb, -ngr.1km)

plot(rast(comb.check, crs = "EPSG:27700", type = "xyz"))
```

```{r}
write_rds(comb, file = here("./Data/Processed_Data/Spatial/map_all_env_vars.rds"))
```






#### Merge env vars with site data to make site.by.env and site.by.xy dataframes
```{r}
# Note, there is sometimes more than one site per gridref
# Merge is used to keep sites that have data for all variables

site.by.env <- sites %>%
  select(site.no, ngr.1km, x, y)
print("Sites with species data")
nrow(sites)

# We need the xy co-ordinates for sites themselves, not their gridsquares
comb <- select(comb, -c(x, y))

site.by.env <- merge(site.by.env, comb, by = "ngr.1km")
nrow(site.by.env)

# Note, there is sometimes more than one site per grid cell

#order by site.number
site.by.env <- site.by.env[base::order(site.by.env$site.no), ]

site.by.env$site <- paste("site", site.by.env$site.no, sep = "_")

# get site as rownamesfor site.by.env
rownames(site.by.env) <- site.by.env$site
site.by.env <- select(site.by.env, -site)

head(site.by.env)
```

#### Plot  Sites
```{r}
plot.points <- st_as_sf(site.by.env, coords=c("x","y"), remove = FALSE, crs = 27700) #I want to keep coords in the dataframe

#get a set of points for each unique grid cell

print("sites")
nrow(plot.points)

basemap <-tm_shape(gb) +
  tm_borders()
pointsmap <- basemap + 
tm_shape(plot.points) +
  tm_symbols(size = 0.05, col = "blue")

pointsmap2 <- basemap + 
tm_shape(plot.points) +
  tm_symbols(size = 0.5, alpha = 0.3, col = "blue")

print(pointsmap)
print(pointsmap2)

```

#### Save as pdf
```{r}
pdf(file= here("./Output/Spatial/env_var_exploration/1122_sites_maps.pdf"))
par(mfrow = c(1,2))
print(pointsmap)
print(pointsmap2)
#turn off PDF plotting
dev.off()
```

## Eyeball environmental data for sampling bias

#### Histograms for environmental data for all grid squares
```{r}

(p1<- ggplot(site.by.temp, aes(x=mean.ann.temp)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "Temperature"))

(p2 <- ggplot(site.by.prec, aes(x=mean.ann.rain)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "Rainfall"))

(p3 <- ggplot(site.by.log10.mean.ann.rain, aes(x=log10.mean.ann.rain)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "Log Rainfall"))

(p4 <- ggplot(site.by.wetness, aes(x=topographic.wetness)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "Topographic Wetness"))

(p5 <- ggplot(site.by.ph, aes(x=ph)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "pH"))


(p6 <- ggplot(site.by.trees, aes(x=tree.density)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "Tree Density"))


(p7 <- ggplot(site.by.nitrogen, aes(x=total.n)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "Total Nitrogen Deposition"))


(p8 <- ggplot(site.by.pesticide.risk, aes(x=pesticide.risk)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "Pesticide Risk"))

(p9 <- ggplot(site.by.log10.pesticide.risk, aes(x=log10.pesticide.risk)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "Log Pesticide Risk"))


(p10 <- ggplot(site.by.humans, aes(x=humans)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "Human Population Density"))


(p11 <- ggplot(site.by.log10.humans, aes(x=log10.humans)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "Log Human Population Density"))

```

#### Histograms for environmental data for sites in dataset
```{r}
(p12 <- ggplot(site.by.env, aes(x=mean.ann.temp)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "Temperature"))

(p13 <- ggplot(site.by.env, aes(x=mean.ann.rain)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "Rainfall"))

(p14 <- ggplot(site.by.env, aes(x=log10.mean.ann.rain)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "Log Rainfall"))

(p15 <- ggplot(site.by.env, aes(x=topographic.wetness)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "Topographic Wetness"))

(p16 <- ggplot(site.by.env, aes(x=ph)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "pH"))

(p17 <- ggplot(site.by.env, aes(x=tree.density)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "Tree Density"))

(p18 <- ggplot(site.by.env, aes(x=total.n)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "Total Nitrogen Deposition"))


(p19 <- ggplot(site.by.env, aes(x=pesticide.risk)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "Pesticide Risk"))
 
(p20 <- ggplot(site.by.env, aes(x=log10.pesticide.risk)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "Log Pesticide Risk"))

(p21 <-ggplot(site.by.env, aes(x=humans)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "Human Population Density"))

(p22 <- ggplot(site.by.env, aes(x=log10.humans)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 20)+
  labs(y = "Probability", title = "Log Human Population Density"))
```

#### Save pdfs of histograms
```{r}
pdf(file= here("./Output/Spatial/env_var_exploration/env_var_histograms_1.pdf"))
figure <- grid.arrange(p1, p2, p3, p4, p12, p13, p14, p15 + rremove("x.text"), 
             ncol = 4, nrow = 2)
dev.off()

pdf(file= here("./Output/Spatial/env_var_exploration/env_var_histograms_2.pdf"))
figure <- grid.arrange(p5, p6, p7, p16, p17, p18 + rremove("x.text"), 
             ncol = 3, nrow = 2)
dev.off()


pdf(file= here("./Output/Spatial/env_var_exploration/env_var_histograms_3.pdf"))
figure <- grid.arrange(p8, p9, p10, p11, p19, p20, p21, p22 + rremove("x.text"), 
             ncol = 4, nrow = 2)
dev.off()

```
#### Check correlation between environmental variables

```{r}
#### Remove untransformed versions of log-transformed variables from environmetal data

site.by.env <- site.by.env %>%
  select(-c(mean.ann.rain, pesticide.risk, humans))

cor(site.by.env[, 3:12], method = "pearson")
```


```{r}
cor(site.by.env[, 3:12], method = "spearman")
```

```{r}
ggpairs(site.by.env, 
        columns = 5:12,
        aes(alpha = 0.05),
        lower = list(continuous = "smooth"))

```


```{r}
ggpairs(site.by.env, 
        columns = 5:12,
        aes(alpha = 0.05),
        lower = list(continuous = "smooth_loess"),
        upper = list(continuous = "smooth_loess"))
```



## Obtain versions of dataset for analysis (i.e. all species, 
## wider countryside species, habitat specialists)
```{r}
# Finalise site.by.xy dataframe for analyses
site.by.xy <- site.by.env %>%
  select(x, y)

# Inspect duplicates for coordinates
dup_xy <- site.by.xy %>%
  mutate(xy = paste(x,y)) %>%
group_by(xy) %>% 
  filter(n() > 1)

# Finalise site.by.env dataframe for analyses
site.by.env <- select(site.by.env, -c(site.no, ngr.1km,  x, y))

rownames(site.by.species) <- paste("site", rownames(site.by.species), sep = "_")
site.by.species <- filter(site.by.species,
                          rownames(site.by.species) %in% rownames(site.by.env))

# Finalise site.by.species dataframe for analyses
print("Number of sites matching butterfly sampling criteria")
nrow(sites)
sites <- sites %>%
  filter(rownames(sites) %in% rownames(site.by.env))
print("Number of sites also with all environmental variables")
nrow(sites)

head(site.by.species)

head(site.by.env)

head(site.by.xy)
```
#### Obtain datasets for wider countryside and habitat specialists
#### With and without rows that have no species
```{r}
species$species.no <- paste("species", species$species.no, sep = "_")

# Obtain vectors of species.no for habitat specialists and wider countryside species
habitat.specialists <- species[species$strategy == "Habitat specialist",]$species.no
wider.countryside.spp <- species[species$strategy == "Wider countryside sp",]$species.no

# Subset site by species by species for habitat specialists and wider countrysides
site.by.species.hs <- site.by.species %>%
  select(all_of(habitat.specialists))

site.by.species.wc <- site.by.species %>%
  select(all_of(wider.countryside.spp))

print("number of sites in all species dataset")
nrow(site.by.species)

print("number of sites with no habitat specialists")
nrow(site.by.species.hs[rowSums(site.by.species.hs)==0, ])

print("number of sites with no wider countryside species")
nrow(site.by.species.wc[rowSums(site.by.species.wc)==0, ])

# filter site by species for wc and hs to remove rows with no species
site.by.species.hs.no.zeros <- site.by.species.hs[rowSums(site.by.species.hs) > 0, ]
print("number of sites with habitat specialists")
nrow(site.by.species.hs.no.zeros)

site.by.species.wc.no.zeros <- site.by.species.wc[rowSums(site.by.species.wc) > 0, ]
print("number of sites with  wider countryside species")
nrow(site.by.species.wc.no.zeros)

wc <- filter_dataset(site.by.species.wc, site.by.xy, site.by.env, species, sites)
hs <- filter_dataset(site.by.species.hs, site.by.xy, site.by.env, species, sites)

wc.no.zeros <- filter_dataset(site.by.species.wc.no.zeros, site.by.xy, site.by.env, species, sites)
hs.no.zeros <- filter_dataset(site.by.species.hs.no.zeros, site.by.xy, site.by.env, species, sites)

toc()
```

#### Save data
```{r}
all_species <- list(site.by.species, site.by.xy, site.by.env, species, sites)
names(all_species) <- c("site.by.species", "site.by.xy", "site.by.env", "species", "sites") 
write_rds(all_species, file = here("./Data/Processed_Data/Spatial/all_species.rds"))

write_rds(wc, file = here("./Data/Processed_Data/Spatial/wc.rds"))
#write_rds(wc.no.zeros, file = here("./Data/Processed_Data/Spatial/wc_no_zeros.rds"))

write_rds(hs, file = here("./Data/Processed_Data/Spatial/hs.rds"))
write_rds(hs.no.zeros, file = here("./Data/Processed_Data/Spatial/hs_no_zeros.rds"))
```
