---
title: "R Notebook"
output: html_notebook
---

#this is copied from ladybirds, I have imported butterfly data but filtering and processing to get all the right size matrices is needed

## setup
```{r}

#The here package enables easy file referencing by using the top-level directory of a file project to easily build file paths."
#about here package: https://here.r-lib.org/articles/here.html
#https://github.com/jennybc/here_here
#by running this before loading the package, we can get a message to check the correct root is being used
here::i_am("Scripts/Rmd_Scripts/butterflies_combine_env_data.Rmd")

library(tidyverse)
library(conflicted)
library(terra)
library(sf)
library(tmap)
library(here)

conflict_prefer("filter", "dplyr", "stats")

# library(readr)
# 
# library(segmented)
# 
# library(sgo)


```


## Load data
```{r}

#load polygon to use to crop as map extents a bit different

gb <- readRDS(here("./Data/Input_Environmental_Data/gb_multipolygon.rds"))

str(gb, max.level = 1)

crs(gb)

#r markdowns and notebooks default to using the directory where they are stored as the working directory


#site by species matrix

# Still need to replace this with occurrence over multiple years
site.by.species <- readRDS(here("./Data/Processed_Data/site_by_species_2015_to_2019.rds"))

#get site by xy
sites <- readRDS(here("./Data/Processed_Data/processed_sites_with_ngr.rds"))

site.by.humans <- readRDS(here("./Data/Input_Environmental_Data/human_popn.rds"))

site.by.humans <- data.frame(site.by.humans)#ideally should have been saved as a df before writing as rds
 
site.by.ph <- readRDS(here("./Data/Input_Environmental_Data/ph_1km_df.rds"))

site.by.trees <-  readRDS(here("./Data/Input_Environmental_Data/trees_1km_df.rds"))

site.by.conifer.dominance <-  readRDS(here("./Data/Input_Environmental_Data/conifer_dominance_1km_df.rds"))

site.by.nitrogen <- readRDS(here("./Data/Input_Environmental_Data/nitro.rds"))

site.by.pesticide.risk <- readRDS(here("./Data/Input_Environmental_Data/pesticide_df.rds"))

site.by.wetness <- readRDS(here("./Data/Input_Environmental_Data/topographic_wetness_df.rds"))

site.by.temp <- readRDS(here("./Data/Input_Environmental_Data/mean_temp_df.rds"))

site.by.prec <- readRDS(here("./Data/Input_Environmental_Data/mean_prec_df.rds"))



surveys.site.year <- readRDS(here("./Data/Processed_Data/surveys_site_year.rds"))

#sites.with.40 <- readRDS(here("./Data/Processed_Data/sites_with_at_least_40_years.rds"))


```

#as a check, turn all input into spatrast and plot
```{r}



# a function to make raster, crop and plot it

#Always does first layer only if multiple layers

plot.df <- function(df, ext){
  rast <- rast(df, crs = "EPSG:27700")
rast <- crop(rast[[1]], ext)
plot(rast)
}


plot.df(df = site.by.temp, ext = gb)


plot.df(df = site.by.prec, ext = gb)

plot.df(df = site.by.wetness, ext = gb)

plot.df(df = site.by.trees, ext = gb)


plot.df(df = site.by.ph, ext = gb)


plot.df(df = site.by.humans, ext = gb)


plot.df(df = site.by.nitrogen, ext = gb)


```



# get a daframe of site by observations, then site by log(observartions)
```{r}

surveys.site.year$site <- paste("site", surveys.site.year$site, sep = "_")




site.by.obs <- surveys.site.year %>%
  ungroup() %>%
    filter(year ==2018)  %>%
      select(-year)  %>%
     rename(observations = surveys)  %>%
      column_to_rownames(var = "site")

site.by.log.obs <- site.by.obs %>%
  mutate(log.observations = log(observations), .keep = "none")



```

#check surveys per site per year
```{r}
surveys.site.year <- surveys.site.year %>%
  filter (year %in% 2015:2019)

hist(surveys.site.year$surveys).

sum <- surveys.site.year %>%
         mutate(surveys = as.character(surveys),
           year = as.character(year)) %>%
  group_by (year, surveys) %>%
  summarise(n = n())



```




## Get log % cover for the sites
```{r}

#keep all vars for now
#site.by.lc <- site.by.lc  %>%
  #  dplyr::select(bro.wood.1, con.wood.2, arable.3, imp.grass.4, seminat.grass.5to8, heath.9to10, freshwater.14, coastal.15to19, #urban.20, suburban.21)

#now we need the transformed land cover composition of the grid cells
logplusone <- function(x)
{log(x + 1)}

#site.by.lc.logs <- site.by.lc %>%
#  mutate(across(colnames(site.by.lc), logplusone)) 
```


## get variable names as desired in dataframes

```{r}
#get site as a variable in the sites and site by species data frames

sites <- rownames_to_column(sites, var = "site")

site.by.species <- rownames_to_column(site.by.species, var = "site")

site.by.ph <-rename(site.by.ph, ph = ph_idw)

site.by.humans <- site.by.humans %>%
  rename(human.density = gbr_pd_2017_1km)%>%
  select(-c(x, y))

site.by.trees <- site.by.trees %>%
  rename(tree.density = tree.cover.density)%>%
  select(-c(x, y)) 

site.by.conifer.dominance <-select(site.by.conifer.dominance, -c(x, y))


site.by.nitrogen <-select(site.by.nitrogen, total.n, ngr.1km)


site.by.pesticide.risk <- site.by.pesticide.risk %>%
  rename(pesticide.risk = pest)%>%
  select(-c(x, y)) 



site.by.wetness <- site.by.wetness %>%
  rename(topgraphic.wetness = Band1) %>%
  select(-c(x,y))

site.by.temp <- site.by.temp %>%
  rename(mean.ann.temp = mean) %>%
  select(-c(x,y))


site.by.prec <- site.by.prec %>%
  rename(mean.ann.rain = mean) %>%
  select(-c(x,y))

```

#filter site data to selected sites

#sometimes more than one site per gridref!
```{r}


#here we filter out sites with anything missing

#filter sites to the desired set 

#sites.2018 <- sites %>%
#     filter(sites$site %in% site.by.species$site)

sites.kept <- sites %>%
     filter(sites$site %in% site.by.species$site)

sites.ngr <- select(sites.kept,site,  ngr.1km)



site.by.env <- merge(sites.ngr, site.by.ph, by = "ngr.1km")


site.by.env <- merge(site.by.env, site.by.humans, by = "ngr.1km") 

site.by.env <- merge(site.by.env, site.by.trees, by = "ngr.1km")




site.by.env <- merge(site.by.env, site.by.nitrogen, by = "ngr.1km")



site.by.env <- merge(site.by.env, site.by.pesticide.risk, by = "ngr.1km")



site.by.env <- merge(site.by.env, site.by.wetness, by = "ngr.1km")


site.by.env <- merge(site.by.env, site.by.temp, by = "ngr.1km")


site.by.env <- merge(site.by.env, site.by.prec, by = "ngr.1km")

```



## Plot  Sites
```{r}

plot.points <- st_as_sf(site.by.env, coords=c("x","y"), remove = FALSE, crs = 27700)#I want to keep coords in the dataframe

#get a set of points for each unique grid cell

print("sites")
nrow(plot.points)

basemap <-tm_shape(gb) +
  tm_borders()
pointsmap <- basemap + 
tm_shape(plot.points) +
  tm_symbols(size = 0.05, col = "blue")

pointsmap2 <- basemap + 
tm_shape(plot.points) +
  tm_symbols(size = 0.5, col = "blue")


print(pointsmap)
print(pointsmap2)


#isle of man and channell islands should probably be excluded
```
#to do:separate WCS and other surveys!
#to do:query sites with too many surveys

```{r}

```



```{r}
hist(site.by.env$mean.ann.temp)
hist(site.by.env$mean.ann.rain)
hist(site.by.env$topgraphic.wetness )#note typo to be corrected
hist(site.by.env$ph)
hist(site.by.env$tree.density)
hist(site.by.env$human.density)
hist(site.by.env$total.n )
hist(site.by.env$pesticide.risk)

```

#transfrom skew variables as spline will work better

```{r}
site.by.env <- site.by.env %>%
    mutate (log.humans = log2(human.density),
          log.pesticide.risk = log2(pesticide.risk +1), 
          .keep = "unused")#remove original columns

#note zeros in pesticide risk!
hist(site.by.env$log.humans)
hist(site.by.env$log.pesticide.risk)
```


```{r}
cor(site.by.env[, 3:12], method = "pearson")
```


```{r}
cor(site.by.env[, 3:12], method = "spearman")


```
#to do: exclude outliers

```{r}

```



```{r}



sites.2018 <- sites.2018 %>%
     filter(sites.2018$ngr.1km %in% site.by.env$ngr.1km)


site.by.species <- site.by.species %>%
     filter(site.by.species$site %in% site.by.env$site)

site.by.species <- column_to_rownames(site.by.species, var = "site")

site.by.env <- column_to_rownames(site.by.env, var = "site")

site.by.xy <- select(site.by.env, x, y)
```



## to do:  change this to the 2015-2019 selection of sites
## look at the percentage of cells that each species occurs in
```{r}
check <- data.frame(colMeans(site.by.species))
  names(check) <- "occurrence"
check <- check %>%
  mutate(var.occurence = occurrence*(1-occurrence))

check
```


```{r}

data.2018 <- list(sites.2018, site.by.species, site.by.env, site.by.xy, site.by.obs)

names(data.2018) <- c("sites.2018", "site.by.species", "site.by.env", "site.by.xy", "site.by.obs")


write_rds(data.2018, file = here("./Data/Processed_Data/butterfly.msgdm.data.2018.rds"))

```

