---
title: "Gam compare splines"
output: html_document

params:
  dataset: "wc"
  family: "gaussian"
---

## In this code, we look at best spline type in GAM, and check 
## variogram for use in GAMM

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Packages
```{r, include=FALSE}
here::i_am("Scripts/Rmd_Scripts/GAM_compare_spline_analyses.Rmd")

library(tidyverse)
library(sf)
library(sp)
library(terra)
library(mgcv)
library(knitr)
library(knitrProgressBar)
library(tmap)
library(mgcViz)
library(gstat)
library(here)
library(conflicted)
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)
conflicts_prefer(knitrProgressBar::progress_estimated)

# there is a same-named function in dplyr, check it out
```

#### Functions
```{r, include=FALSE}
source(here("Scripts/R_scripts/butterfly_richness_turnover_functions.R"))
```

#### File Paths
```{r, include=FALSE}
data_path <- paste0("./Data/Processed_Data/Spatial/", params$dataset, ".rds")

out_path <-  paste0(here("Output/Spatial/GAM"), "/","GAM", "_" , 
                    params$dataset,  "/")

if(!dir.exists(out_path)) dir.create(out_path)
```

#### Data
```{r}
data <- readRDS(here(data_path))

site.by.species <- data$site.by.species
  
site.by.env <- data$site.by.env

site.by.xy <- data$site.by.xy

# gb polygon

gb <- readRDS(here("./Data/Maps/gb_multipolygon_simplified.rds"))
```

```{r}
print(params)
```
#### Prepare data for analysis
```{r}
site.by.env.and.richness <- cbind(site.by.xy) %>%
  cbind(site.by.env, rowSums(site.by.species)) %>% 
  rename(species.richness = 'rowSums(site.by.species)')
```

## Run GAMs

#### Thin plate splines
```{r}
gam_tp <-  gam(species.richness ~ s(mean.ann.temp) + s(log10.mean.ann.rain) + 
                s(topographic.wetness) + s(ph) + s(total.n) + s(tree.density) + 
                s(log10.pesticide.risk) + s(log10.humans),  
                data = site.by.env.and.richness, family = params$family, method= "REML")
```

#### Cubic regression splines
```{r}
gam_cr <- gam(species.richness ~ s(mean.ann.temp, bs="cr") + 
            s(log10.mean.ann.rain, bs="cr") + s(topographic.wetness, bs="cr") + 
            s(ph, bs="cr") + s(total.n, bs="cr") + s(tree.density, bs="cr") + 
            s(log10.pesticide.risk, bs="cr") + s(log10.humans, bs="cr"),  
                data = site.by.env.and.richness, family = params$family, method= "REML")
```

#### compare AIC 
```{r}
gam_list <-list (gam_tp,  gam_cr)
names(gam_list) <- c("gam_tp",  "gam_cr")
# 
# aic_table <- AIC(gam_tp,  gam_cr) %>%
#   arrange(AIC) %>%
#   mutate(delta_AIC = AIC - min(AIC)) %>%
#   rownames_to_column(var = "Splines")
# 
# kable(aic_table, digits = 2, align = "l")
# 
# write_csv(aic_table, file = paste0(out_path, "/", "spline_aic_table.csv"))
```

### Compare residual plots and perform basis dimension checking
```{r}
for (i in 1:length(gam_list)){
  print(names(gam_list[i]))  
  gam_i <- gam_list[[i]]
  par(mfrow = c(2, 2))
  gam.check(gam_i)
}
```
### Compare GAM summaries
```{r}
for (i in 1:length(gam_list)){
  print(names(gam_list[i]))  
  gam_i <- gam_list[[i]]
  print(summary(gam_i))
}
```

### Compare GAM smooth plots
```{r}
for (i in 1:length(gam_list)){
  print(names(gam_list[i]))  
  gam_i <- gam_list[[i]]
  par(mfrow = c(2, 2))
  my.plot.gam(gam_i)
}
```

### Compare Moran's tests
```{r}
for (i in 1:length(gam_list)){
  print(names(gam_list[i]))  
  gam_i <- gam_list[[i]]
  print("Moran's test  with I value")
  print(moranfast(residuals(gam_i, type = "deviance"), 
          site.by.env.and.richness$x, site.by.env.and.richness$y))
}
```

#### Geographic pattern in residuals
```{r}
for (i in 1:length(gam_list)){
  print(names(gam_list[i]))  
  gam_i <- gam_list[[i]]
  
residuals <- resid(gam_i)
xy.res <- cbind(site.by.xy, residuals)


v1 <-   variogram(residuals ~ 1, loc= ~ x+y, data = xy.res)
fit1 <- fit.variogram(v1, vgm(c("Sph", "Exp"))); 
print(plot(v1, fit1))

filepath <- paste0(out_path, "/variogram_", names(gam_list[i]), ".png")
unlink(filepath, recursive = FALSE, force = FALSE) #deletes previous version with the same name

# site.by.env.and.richness <- cbind(site.by.xy) %>%
#   cbind(site.by.env, rowSums(site.by.species)) %>% 
#   rename(species.richness = 'rowSums(site.by.species)')

env.and.res <- cbind(site.by.env.and.richness, residuals)

coordinates(xy.res) <- c("x","y")

crs.geo <- CRS("+init=epsg:27700")  # looks up BNG crs
proj4string(xy.res) <- crs.geo  # define projection system of our data

print(bubble(xy.res, "residuals", col = c("black","grey"),
       main = "Residuals", xlab = "X-coordinates", ylab = "Y-coordinates"))

basemap <-tm_shape(gb) +
  tm_borders()

pointsmap <- basemap +
tm_shape(xy.res) +
  tm_symbols(size = 0.25, col = "residuals", midpoint = NA)
print(pointsmap)


png(file = filepath)
plot(v1, fit1)
dev.off()
}
```

### Compare GAMs in terms of AIC
```{r}
aic_table <- AIC(gam_tp, gam_cr) %>%
  arrange(AIC) %>% #order rows by AIC
  mutate(delta_AIC = AIC - min(AIC)) %>%
  rownames_to_column(var = "Family")

write_csv(aic_table, file = paste0(out_path, "/", "spline_aic_table.csv"))

kable(aic_table, digits = 2, align = "l")
```