---
title: "GAMMs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Packages
```{r}
here::i_am("Scripts/Rmd_Scripts/GAMM_analyses.Rmd")
library(tidyverse)
library(tmap)
library(mgcv)
library(gstat)
library(knitr)
library(knitrProgressBar)
library(tictoc)
library(here)
library(conflicted)
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)
conflicts_prefer(knitrProgressBar::progress_estimated)
# there is a same-named function in dplyr, check it out
```

#### Functions
```{r}
# From https://stackoverflow.com/questions/71252509/is-there-a-way-to-get-progress-information-whan-a-chunk-is-running-in-rmarkdown

slow_function <- function(i, .pb=NULL) {  
  update_progress(.pb)  
  Sys.sleep(0.5)  
  i  
}
# create an R6 progress object for
# the number of loops/iterations in the target chunk
```

#### File Paths
```{r}
out_path <-  here("Output/Spatial/GAM")
if(!dir.exists(out_path)) dir.create(out_path)
```

#### Datasets for gamms
```{r}

all_species <- readRDS(here("./Data/Processed_Data/Spatial/all_species.rds"))
wc <- readRDS(here("./Data/Processed_Data/Spatial/wc.rds"))
hs <- readRDS(here("./Data/Processed_Data/Spatial/hs.rds"))
hs_no_zeros <- readRDS(here("./Data/Processed_Data/Spatial/hs_no_zeros.rds"))

# get all datasets in a list
datasets <- list(all_species, wc, hs, hs_no_zeros)
names(datasets) <- c("all_species", "wc", "hs", "hs_no_zeros")
```



#### for each dataset, prepare data for gamm input, remove transects that are duplicates for x and y coordinates (keep first one)
```{r}
for (i in 1: length(datasets)){ 
print(names(datasets[i]))
  
data <- datasets[[i]]

site.by.species <- data$site.by.species
site.by.env <- data$site.by.env
site.by.xy <- data$site.by.xy
sites <- data$sites

# assemble xy, environmental vars and richness for analysis
site.by.env.xy.richness <- cbind(site.by.env, rowSums(site.by.species)) %>%
  cbind(site.by.xy) %>%
rename(species.richness = 'rowSums(site.by.species)')

#inspect duplicates to see how they arise
dup_xy <- sites %>%
  mutate(xy = paste(x,y)) %>%
group_by(xy) %>% 
  filter(n() > 1)

print(kable(dup_xy, caption = names(datasets[i])))
  
#remove duplicates for rows with the same x and y coordinates
gamm_input <- site.by.env.xy.richness %>%
  mutate(xy = paste(x,y)) %>%
  distinct(xy, .keep_all = TRUE)

hist(gamm_input$species.richness, main = names(datasets[i]))

# add the input for the gamm to each dataset
write_rds(gamm_input, 
          file = here(
            paste0("./Data/Processed_Data/Spatial/", names(datasets[i]), "_gamm_input.rds" )
            )
          )

gamm_input <- list(gamm_input)
names(gamm_input) <- "gamm_input"
datasets[[i]] <- c(datasets[[i]], gamm_input)
}


```

#### Specify  correlation structures
```{r}
corStructs <- list(corGaus(form = ~x+y, nugget = TRUE), corExp(form = ~x+y, nugget = TRUE), corLin(form = ~x+y, nugget = TRUE), corRatio(form = ~x+y, nugget = TRUE), corSpher(form = ~x+y, nugget = TRUE) )
```


## run GAMMs and save


# Full model before dropping terms based on concurvity


### Annoyingly, I did not succeed in a wrapper function and in the Markdown version, it was hanging
### Therefore, I have copied and pasted and checked carefully


#### All species, thin plate splines, gaussian family
```{r}

#check list to modify after copy past

#formula
#family
#data
#filedir

gamm_list <- list()
# Loop through the gamms saving output in list
# Nugget allows non-zero autocorrelation at 0 distance
for (i in 1:length(corStructs)) {
my_gamm <- gamm(
  species.richness ~ s(mean.ann.temp, k = 5) + s(log10.mean.ann.rain, k = 5) + 
  s(topographic.wetness, k = 5) + s(ph, k = 5) + s(total.n, k = 5) + s(tree.density, k = 5) + 
  s(log10.pesticide.risk, k = 5) + s(log10.humans, k = 5), 
  method = "REML", 
  family = "gaussian", 
  correlation = corStructs[[i]], 
  data = datasets$all_species$gamm_input)

print(i)
print(Sys.time())

gamm_list[[i]] <- my_gamm

}

filedir <- paste0(here(out_path), "/GAMM_all_species")
if(!dir.exists(filedir)) dir.create(filedir)
filepath <- paste0(filedir, "/gamm_cor_struct_compare.rds")
filepath
write_rds(gamm_list , 
          file = filepath)
```

#### Wider countryside, thin plate splines, gaussian family
```{r}

#checklist to modify after copy past

#formula
#family
#data
#filedir

gamm_list <- list()
# Loop through the gamms saving output in list
# Nugget allows non-zero autocorrelation at 0 distance
for (i in 1:length(corStructs)) {
my_gamm <- gamm(
  species.richness ~ s(mean.ann.temp, k = 5) + s(log10.mean.ann.rain, k = 5) + 
  s(topographic.wetness, k = 5) + s(ph, k = 5) + s(total.n, k = 5) + s(tree.density, k = 5) + 
  s(log10.pesticide.risk, k = 5) + s(log10.humans, k = 5), 
  method = "REML", 
  family = "gaussian", 
  correlation = corStructs[[i]], 
  data = datasets$wc$gamm_input)

print(i)
print(Sys.time())

gamm_list[[i]] <- my_gamm

}

filedir <- paste0(here(out_path), "/GAMM_wc")
if(!dir.exists(filedir)) dir.create(filedir)
filepath <- paste0(filedir, "/gamm_cor_struct_compare.rds")
filepath
write_rds(gamm_list , 
          file = filepath)
```


#### Habitat specialist, thin plate splines, nb
```{r}
#checklist to modify after copy past

#formula
#family
#data
#filedir
#write_rds

gamm_list <- list()
# Loop through the gamms saving output in list
# Nugget allows non-zero autocorrelation at 0 distance
for (i in 1:length(corStructs)) {
my_gamm <- gamm(
  species.richness ~ s(mean.ann.temp, k = 5) + s(log10.mean.ann.rain, k = 5) + 
  s(topographic.wetness, k = 5) + s(ph, k = 5) + s(total.n, k = 5) + s(tree.density, k = 5) + 
  s(log10.pesticide.risk, k = 5) + s(log10.humans, k = 5), 
  method = "REML", 
  family = negbin(theta = 4.679), 
  correlation = corStructs[[i]], 
  data = datasets$hs$gamm_input)

print(i)
print(Sys.time())

gamm_list[[i]] <- my_gamm

}

filedir <- paste0(here(out_path), "/GAMM_hs")
if(!dir.exists(filedir)) dir.create(filedir)
filepath <- paste0(filedir, "/gamm_cor_struct_compare.rds")
filepath
write_rds(gamm_list , 
          file = filepath)
```



# Reduced model without rainfall, pesticide, trees to reduce concurvity

### All Species
```{r}
gamm_as <- gamm(
  species.richness ~ s(mean.ann.temp, k = 5) +
  s(topographic.wetness, k = 5) + s(ph, k = 5) + s(total.n, k = 5) 
  + s(log10.humans, k = 5), 
  method = "REML", 
  correlation = corStructs[[2]], #corExp
  data = datasets$all_species$gamm_input)


filedir <- paste0(here(out_path), "/GAMM_all_species")
if(!dir.exists(filedir)) dir.create(filedir)
filepath <- paste0(filedir, "/gamm_all_species_exp_gaus2_reduced_mod.rds")
filepath
write_rds(gamm_as, 
          file = filepath)
```

#### All species, thin plate splines, gaussian family
```{r}

#check list to modify after copy past

#formula
#family
#data
#filedir

gamm_list <- list()
# Loop through the gamms saving output in list
# Nugget allows non-zero autocorrelation at 0 distance
for (i in 1:length(corStructs)) {
my_gamm <- gamm(
  species.richness ~ s(mean.ann.temp, k = 5) + s(log10.mean.ann.rain, k = 5) + 
  s(topographic.wetness, k = 5) + s(ph, k = 5) + s(total.n, k = 5) + s(tree.density, k = 5) + 
  s(log10.pesticide.risk, k = 5) + s(log10.humans, k = 5), 
  method = "REML", 
  family = "gaussian", 
  correlation = corStructs[[i]], 
  data = datasets$all_species$gamm_input)

print(i)
print(Sys.time())

gamm_list[[i]] <- my_gamm

}

filedir <- paste0(here(out_path), "/GAMM_all_species")
if(!dir.exists(filedir)) dir.create(filedir)
filepath <- paste0(filedir, "/gamm_cor_struct_compare.rds")
filepath
write_rds(gamm_list , 
          file = filepath)
```

#### Wider countryside, thin plate splines, gaussian family
```{r}

#checklist to modify after copy past

#formula
#family
#data
#filedir

gamm_list <- list()
# Loop through the gamms saving output in list
# Nugget allows non-zero autocorrelation at 0 distance
for (i in 1:length(corStructs)) {
my_gamm <- gamm(
  species.richness ~ s(mean.ann.temp, k = 5) + s(log10.mean.ann.rain, k = 5) + 
  s(topographic.wetness, k = 5) + s(ph, k = 5) + s(total.n, k = 5) + s(tree.density, k = 5) + 
  s(log10.pesticide.risk, k = 5) + s(log10.humans, k = 5), 
  method = "REML", 
  family = "gaussian", 
  correlation = corStructs[[i]], 
  data = datasets$wc$gamm_input)

print(i)
print(Sys.time())

gamm_list[[i]] <- my_gamm

}

filedir <- paste0(here(out_path), "/GAMM_wc")
if(!dir.exists(filedir)) dir.create(filedir)
filepath <- paste0(filedir, "/gamm_cor_struct_compare.rds")
filepath
write_rds(gamm_list , 
          file = filepath)
```


#### Habitat specialist, thin plate splines, nb
```{r}
#checklist to modify after copy past

#formula
#family
#data
#filedir
#write_rds

gamm_list <- list()
# Loop through the gamms saving output in list
# Nugget allows non-zero autocorrelation at 0 distance
for (i in 1:length(corStructs)) {
my_gamm <- gamm(
  species.richness ~ s(mean.ann.temp, k = 5) + s(log10.mean.ann.rain, k = 5) + 
  s(topographic.wetness, k = 5) + s(ph, k = 5) + s(total.n, k = 5) + s(tree.density, k = 5) + 
  s(log10.pesticide.risk, k = 5) + s(log10.humans, k = 5), 
  method = "REML", 
  family = negbin(theta = 4.679), 
  correlation = corStructs[[i]], 
  data = datasets$hs$gamm_input)

print(i)
print(Sys.time())

gamm_list[[i]] <- my_gamm

}

filedir <- paste0(here(out_path), "/GAMM_hs")
if(!dir.exists(filedir)) dir.create(filedir)
filepath <- paste0(filedir, "/gamm_cor_struct_compare.rds")
filepath
write_rds(gamm_list , 
          file = filepath)
```



# Reduced model without rainfall, pesticide, trees to reduce concurvity



#### All species, thin plate splines, gaussian family
```{r}

#check list to modify after copy past

#formula
#family
#data
#filedir

gamm_list <- list()
# Loop through the gamms saving output in list
# Nugget allows non-zero autocorrelation at 0 distance
for (i in 1:length(corStructs)) {
my_gamm <- gamm(
  species.richness ~ s(mean.ann.temp, k = 5) + 
  s(topographic.wetness, k = 5) + s(ph, k = 5) + s(total.n, k = 5) + 
  s(log10.humans, k = 5), 
  method = "REML", 
  family = "gaussian", 
  correlation = corStructs[[i]], 
  data = datasets$all_species$gamm_input)

print(i)
print(Sys.time())

gamm_list[[i]] <- my_gamm

}

filedir <- paste0(here(out_path), "/GAMM_all_species")
if(!dir.exists(filedir)) dir.create(filedir)
filepath <- paste0(filedir, "/gamm_cor_struct_compare_reduced_mod.rds")
filepath
write_rds(gamm_list , 
          file = filepath)
```

#### Wider countryside, thin plate splines, gaussian family
```{r}

#checklist to modify after copy past

#formula
#family
#data
#filedir

gamm_list <- list()
# Loop through the gamms saving output in list
# Nugget allows non-zero autocorrelation at 0 distance
for (i in 1:length(corStructs)) {
my_gamm <- gamm(
  species.richness ~ s(mean.ann.temp, k = 5) + 
  s(topographic.wetness, k = 5) + s(ph, k = 5) + s(total.n, k = 5) + 
  s(log10.humans, k = 5), 
  method = "REML", 
  family = "gaussian", 
  correlation = corStructs[[i]], 
  data = datasets$wc$gamm_input)

print(i)
print(Sys.time())

gamm_list[[i]] <- my_gamm

}

filedir <- paste0(here(out_path), "/GAMM_wc")
if(!dir.exists(filedir)) dir.create(filedir)
filepath <- paste0(filedir, "/gamm_cor_struct_compare_reduced_mod.rds")
filepath
write_rds(gamm_list , 
          file = filepath)
```


#### Habitat specialist, thin plate splines, nb
```{r}
#checklist to modify after copy past

#formula
#family
#data
#filedir
#write_rds

gamm_list <- list()
# Loop through the gamms saving output in list
# Nugget allows non-zero autocorrelation at 0 distance
for (i in 1:length(corStructs)) {
my_gamm <- gamm(
  species.richness ~ s(mean.ann.temp, k = 5) + 
  s(topographic.wetness, k = 5) + s(ph, k = 5) + s(total.n, k = 5) + 
  s(log10.humans, k = 5), 
  method = "REML", 
  family = negbin(theta = 4.679), 
  correlation = corStructs[[i]], 
  data = datasets$hs$gamm_input)

print(i)
print(Sys.time())

gamm_list[[i]] <- my_gamm

}

filedir <- paste0(here(out_path), "/GAMM_hs")
if(!dir.exists(filedir)) dir.create(filedir)
filepath <- paste0(filedir, "/gamm_cor_struct_compare_reduced_mod.rds")
filepath
write_rds(gamm_list , 
          file = filepath)
```
