---
title: "R Notebook"
output: html_notebook
---
`
# This makes figure 2

```{r}
here::i_am("Scripts/Rmd_Scripts/Gamm_multipanel_fig.Rmd")
library(tidyverse)
library(gridExtra)
library(ggpubr)
library(mgcv)
library(mgcViz)
library(here)
library(conflicted)

conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)
```

#### Functions
```{r}
# A function to prepare data from plot.gam() and prepare for further plotting.
# plot_data is a list with length being the number of predictors in the GAM.

prepare.clfit <- function(shift, plot_data){

  outlist <- list()
 
for (i in 1:length(plot_data)){
  input <- plot_data[[i]]
  output <-  data.frame(cbind(
  input$x, 
  input$fit + shift, 
  input$se)) #note, se already multiplied by 1.96
  names(output) <- c("x", "fit", "se.fit")
  output <- output %>%
  mutate(hi.cl = fit + se.fit, lo.cl = fit - se.fit)
  outlist[[i]] <- output
  } 
return(outlist)  
}

# a function to plot fit and ci from plot.gam
plot.clfit <- function(df){
df %>%
ggplot(aes(x)) +
geom_line(aes(y = fit), colour = "red", linetype = 1) +
geom_line(aes(y = hi.cl), colour = "blue", linetype = 2) +
geom_line(aes(y = lo.cl), colour = "blue", linetype = 2)  + 
# geom_rug(sides="b")  + 
  theme_classic()  
}

```

#### File paths
```{r}
out_path <-  here("Output/Spatial/GAM/GAMM_fig//")
if(!dir.exists(out_path)) dir.create(out_path)
```

#### Load gamm objects
```{r}
# they are in this order: ("gamm_corGaus", "gamm_corExp", "gamm_corLin",
#"gamm_corRatio", "gamm_corSpher")

as_mods <- readRDS(here("./Output/Spatial/GAM/GAMM_all_species/gamm_model_select_as.rds/"))
as_gamm <- as_mods[[3]]

wc_mods <- readRDS(here("./Output/Spatial/GAM/GAMM_wc/gamm_model_select_wc.rds/"))
wc_gamm <- wc_mods[[3]]

hs_mods <- readRDS(here("./Output/Spatial/GAM/GAMM_hs/gamm_model_select_hs.rds/"))
hs_gamm <- hs_mods[[3]]
```

#### Get predicted values and SE for smooths from gam.plot and inspect plots provided by mgcv
```{r}

as_plot_data <- plot(as_gamm$gam, shift = coef(as_gamm$gam)[1], pages = 1, seWithMean = FALSE, se = 1.96) # in the output se is already multiplied by 1.96

wc_plot_data <- plot(wc_gamm$gam, shift = coef(wc_gamm$gam)[1], pages = 1, seWithMean = FALSE, se = 1.96) # in the output se is already multiplied by 1.96

hs_plot_data <- plot(hs_gamm$gam, shift = coef(hs_gamm$gam)[1], pages = 1, seWithMean = FALSE, se = 1.96) # in the output se is already multiplied by 1.96

```

#### prepare data from gam.plot to make a custom plot with geom_line()
```{r}
as_plot_data2 <- prepare.clfit(coef(as_gamm$gam)[1], as_plot_data)
wc_plot_data2 <- prepare.clfit(coef(wc_gamm$gam)[1], wc_plot_data)
hs_plot_data2 <- prepare.clfit(coef(hs_gamm$gam)[1], hs_plot_data)
```

#### create plots for gam splines and add to a list
```{r}
my_panels <- list()

for (i in 1:6){
my_panels[[i]] <- plot.clfit(as_plot_data2[[i]])
}

for (i in 7:12){
my_panels[[i]] <- plot.clfit(wc_plot_data2[[i- 6]])
}

for (i in 13:18){
my_panels[[i]] <- plot.clfit(hs_plot_data2[[i - 12]])
}
```

#### Modify panels by adding labels (letters) and specifying axis titles
```{r}
# Column headings
top_heads <- c("All", "Generalists", "Specialists")
# Panel labels
panel_letters <- c("A", "D", "G", "J", "M", "P", 
                   "B", "E", "H", "K", "N", "Q", 
                   "C", "F", "I", "L", "O", "R")
# x-axis legends
x_labels <- c("Temperature", "Log Prec.", "Top. Wet.", "pH", "Nit. Dep.", "Log Humans")


for (i in 1:18){
  # for each panel adjust appearance as follows
  p_i <-  my_panels[[i]]
  # Remove x axis legend for panels not in middle column and rename legend for panels that are in the middle column
  if (!is.element(i, 7:12)) {
    p_i <-  p_i + 
      xlab(" ")
    } else {
    p_i <- p_i + 
      xlab(x_labels[i - 6])    
    }
  # only include y axis legend for the middle left panel
  if (i == 3) {
    p_i <-  p_i + 
      ylab("Species Richness") 
  } else {
    p_i <-  p_i + 
      ylab(" ")
  }
    # add panel labels to each panel
  p_i <- p_i +
    annotation_custom(grid::textGrob(panel_letters[i], 0.15, 0.9, gp = grid::gpar(fontsize = 6))) 
  my_panels[[i]] <- p_i
}

```

#### create a new list with the panels in the order needed for plottin (they are plotted row-wize but my_panels has them column_Wize)
```{r}
# my_index gives index in the original list of panels for new list, but in the order the should be in for the new list. 
# start_index is the first three entries for my_index, subsequent values of my_index are obtained by adding 1, 2, 3... to start_index entries
start_index <- c(1, 7, 13)
my_index <- start_index
for (i in 1:5){
  my_index <- c(my_index, start_index +i)
}
my_panels2 <- list()

for (i in 1:27){
  my_panels2[[i]]<- my_panels[[my_index[i]]]
}
```

#### Make and inspect multipanel plot
```{r}
nr <- 6
nc <- 3

# to include column caption as a separate row, I addapted code from https://stackoverflow.com/questions/45473843/put-row-and-column-titles-using-grid-arrange-in-r
all_elements <- rbind(tableGrob(t(top_heads), theme = ttheme_minimal(), rows = ""), 
                 cbind(tableGrob(rep(" ",nr), theme = ttheme_minimal()), # I couldn't work out how not to have row headings so made them a space 
                       arrangeGrob(grobs = my_panels2, ncol = nc),  size = "last"), size = "last")
grid.newpage()
grid.draw(all_elements)
```
#### write multipanel plot to png
```{r}
png(file = paste0(out_path, "gam_multiplot.png"), pointsize=10, width=1920, height=2800, res=300)
grid.newpage()
grid.draw(all_elements)
dev.off()

```

#### example code from  https://stackoverflow.com/questions/45473843/put-row-and-column-titles-using-grid-arrange-in-r
```{r}
pl <- replicate(12, ggplot(), FALSE)

#N <- length(pl)
nr <- 4
nc <- 3


all_elements <- rbind(tableGrob(t(c(letters[1:nc])), theme = ttheme_minimal(), rows = ""), 
                 cbind(tableGrob(LETTERS[1:nr], theme = ttheme_minimal()), 
                       arrangeGrob(grobs = pl),  size = "last"), size = "last")
grid.newpage()
grid.draw(all_elements)
```