---
title: "R Notebook"
output: html_notebook
---

#### datasets used for generating results
```{r}
datasets <- c("all_species", "wc", "hs", "hs_no_zeros")
datasets_names <- c("All Butterfly Species", "Wider Countryside Species", "Habitat Specialists including sites with no species", "Habitat Specialists excluding sites with no species" )
datasets2 <- c("all_species", "wc")
datasets_names2 <- c("All Butterfly Species", "Wider Countryside Species")
```

#### make chunks for zeta decline figs 1:10
```{r}
for (i in 1:length(datasets)) {
  cat(
    paste0("#### ", datasets_names[i],
           "\n",
          '```{r fig.cap = "Figure S',i+2, ". Zeta diversity decline for ", datasets_names[i], ', zeta orders 1 to 10", out.width = ','"100%"}',
          "\n",
          'include_graphics(here("./Output/Spatial/zeta_declines_decays/', datasets[i], "/", datasets[i], '_zeta_decline_to_10.png"))',
          "\n",
          "```",
          "\n\n"
           )
  )
}
```
#### make chunks for zeta decline figs 1:250

```{r}
for (i in 1:length(datasets2)) {
  cat(
    paste0("#### ", datasets_names2[i],
           "\n",
          '```{r fig.cap = "Figure S',i+6, ". Zeta diversity decline for ", datasets_names2[i], ', zeta orders 1 to 10", out.width = ','"100%"}',
          "\n",
          'include_graphics(here("./Output/Spatial/zeta_declines_decays/', datasets2[i], "/", datasets2[i], '_zeta_decline_to_250.png"))',
          "\n",
          "```",
          "\n\n"
           )
  )
}
```