---
title: "Process_butterflies"
output: html_notebook
---



## Data cleaning, obtaining site by species matrices
```{r}

here::i_am("Scripts//Rmd_Scripts//Process_CSV_data.Rmd")

library(conflicted)
library(tidyverse)
library(here)

#library(foreach)
#library(doParallel)
#library(doSNOW)

library(spaa)

library(sgo)

library(sf)

library(tidyterra)

library(scales)

library(segmented)

library(tmap)

library(Hmisc)


library(iNEXT)

conflicts_prefer(snow::stopCluster)
conflicts_prefer(snow::makeCluster)
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::mutate)
conflicts_prefer(dplyr::select)
#filter <- dplyr::filter


```


```{r}
#(cl = makeCluster(parallel::detectCores()-2, type = "SOCK")) # Initialize all cores except for two
#registerDoSNOW(cl)


records <- read_csv(here("./Data/CSV_Data/ukbms_Rogers_10-3-23.csv"))

#In the CSV, I  added in column headings, did a sort and added in extra columns to help correct entries in incorrect columns
#note, some sites have incomplete info, one site remains with latitude and longitude in wrong columns and no other info, such sites are filtered in this file


sites <- read_csv(here("./Data/CSV_Data/ukbms_Rogers_sites_10-3-23_to_correct.csv"))

sites.more.info <- read_csv(here("./Data/CSV_Data/ukbmsSiteLocationData1976-2019.csv"))

species <- read_csv(here("./Data/CSV_Data/ukbms_Rogers_species_10-3-23.csv"))

bounds <- readRDS(here("./Data./Input_Environmental_Data./gb_multipolygon_incl_ork_shet.rds"))

bounds.2 <- readRDS(here("./Data./Input_Environmental_Data./gb_multipolygon.rds"))

```

```{r}
head(records)
summary(records$YEAR)
```
## this shows that column names need attention

## inspecting the dataframe shows some info is in the wrong columns
```{r}
head(sites)
nrow(sites)
```

## clean and organise site data
```{r}

#rename columns
names(sites)
sites <- sites %>%
  rename(site = SITENO,
         site.name =  SITENAME,
         gridref = GRIDREF,
         longitude = LONGITUDE,
         latitude = LATITUDE) %>% 
  #here we move values that are in the wrong columns to the correct columns
  mutate(site.name = case_when(paste2cols == 1 ~ paste(site.name, gridref, sep = "_"),
                             paste3cols == 1 ~ paste(site.name, gridref, longitude, sep = "_"),
                             .default = as.character(site.name)),
         gridref = case_when(paste2cols == 1 ~ as.character(longitude),
                             paste3cols == 1 ~ as.character(latitude),
                             .default = as.character(gridref)),
         longitude = case_when(paste2cols == 1 ~ as.numeric(latitude),
                             paste3cols == 1 ~ as.numeric(extracol1),
                             .default = as.numeric(longitude)), 
         latitude = case_when(paste2cols == 1 ~ as.numeric(extracol1),
                             paste3cols == 1 ~ as.numeric(extracol2),
                             .default = as.numeric(latitude))) %>% 
  dplyr::select(-c(extracol1, extracol2, paste2cols, paste3cols))


# for inspection only, sites that have no co-ordinates or grid square reference


bad.sites <- filter(sites, is.na(longitude) == T &  is.na(gridref) ==T)

print("number of sites with inadequate locality info")
nrow(bad.sites)

#remove irish sites, which have uneven number of characters in the gridref (they start with a single letter)
sites <- sites %>%
                mutate(str.len.gridref = str_length(gridref),# a variable showing number of characters in gridref
                       even = str.len.gridref %% 2 == 0)  #whether above variable is even


irish.sites <- filter(sites, even == F)  #for inspection only


print("number of Northern Irish sites")
nrow(irish.sites)


print("number of sites before filterin out ones without locality info and ones in Ireland")
nrow(sites)

#can be used
sites <-  filter(sites, is.na(longitude) == F |  is.na(gridref) == F,#remove 51 sites with neither a gridref nor longlat
                even == T) %>% #filter 138 Irish sites
                    dplyr::select(-c(even, str.len.gridref))

#one wonky site not picked up above
sites <- filter(sites, gridref != "Lurgan") 


print("number of sites after filtering out ones without locality info and ones in Ireland")
nrow(sites)     
```


```{r}
dplyr::last_dplyr_warnings()
head(species)
nrow(species)
```
#add in missing longitude and latitude for 20 sites
```{r}

#obtain longitude and latitude using sgo

#just work with rows with missing dtaa
for.longlat <- sites %>% 
  filter(is.na(longitude) == TRUE)

#get eastings and northings first
for.longlat <- sgo_ngr_bng(for.longlat, col = "gridref")

#get latitude and longitude
for.longlat <- sgo_bng_lonlat(for.longlat)

#get a dataframe of rows with missing data to add them into
addin <-sites %>%
filter(is.na(longitude) == TRUE) %>%
  rename(long = longitude, 
         lat = latitude)

#add in longitude and latitude
addin$long <- for.longlat$x


addin$lat <- for.longlat$y


addin <- dplyr::select(addin,
                site, long, lat)

#merge with the sites data

print("number of sites before merging operation")
nrow(sites)
sites <- merge(sites, all = T, addin, by = "site")

print("number of sites after merging operation")
nrow(sites)

#put data into the correct column and remove now extraneous columns

sites <- sites %>%
  mutate(
    longitude = ifelse(is.na(longitude) == TRUE, long, longitude),
    latitude = ifelse(is.na(latitude) == TRUE, lat, latitude)
  ) %>%
  dplyr::select(-c(lat, long))

#get easting and wesing for points with latitude and longitude
pts <- sgo_points(sites, coords = c("longitude", "latitude"), epsg=4277)#this epsg for long and lat
bng.pts1 <- sgo_lonlat_bng(pts)


pts.df1 <- as.data.frame(bng.pts1)

pts.df1 <- pts.df1 %>% 
  rename(
       easting = x,
       northing = y) %>% 
  dplyr::select(easting, northing, site)
  


# put together eastings 

print("number of sites before merging operation")
nrow(sites)
sites<- merge(sites, pts.df1, by = "site")
 print("number of sites after merging operation")
 nrow(sites)
```




## Plot Sites
```{r}
# 
# plot.points <- st_as_sf(sites, coords=c("easting","northing"), remove = FALSE, crs = 27700)#I want to keep coords in the dataframe
# 
# #get a set of points for each unique grid cell
# 
# print("sites")
# nrow(plot.points)
# 
# basemap <-tm_shape(bounds) +
#   tm_borders()
# pointsmap <- basemap + 
# tm_shape(plot.points) +
#   tm_symbols(size = 0.05, col = "blue")
# 
# pointsmap2 <- basemap + 
# tm_shape(plot.points) +
#   tm_symbols(size = 0.5, col = "blue")
# 
# 
# print(pointsmap)
# print(pointsmap2)
```


### filter species and records to exclude unknown species and moths, as well as surveys where no butterflies were seen,  and migrant, vagrant, extinct species


### filter sites to those included in records

```{r}

#rename columns in species and...


print("number of species in original species sheet")
nrow(species)

species <- species %>% 
  rename( species.no = SPECIES, 
         common.name = COMMON_NAME,
         sci.name = SCI_NAME,
         strategy = STRATEGY,
         habitat = HABITAT) %>% 
  filter (is.na(include) == FALSE,
          strategy == "Wider countryside sp"| strategy == "Habitat specialist") #filter out non-butterflies and "unidentified" from species list

print("number of species after removing moths, migrants, and unidentified spp")
nrow(species)

#so we can use it for summarising data etc later
species <- species %>% 
  mutate(species.no = as.character(species.no))

#rename columns in records
records<- records %>%
    rename(site = SITE, 
        species.no = SPECIES,
        year = YEAR,
        month = MONTH, 
        day = DAY, 
        week.no = WEEKNO, 
        day.no = DAYNO, 
        count = COUNT)%>% 
   mutate(species.no = as.character(species.no))

print("rows of records")
nrow(records)

print("number of records for all species including moths and migrant butterflies")
nrow(records)

records <- records %>% 
  filter (species.no %in% species$species.no) #this filters out records that were not butterflies or were not identified
      

print("number of records of butterflies excluding migrants")
nrow(records)



records <- records %>% 
  filter ( site %in% sites$site)


print("number of records for remaining butterflies  in remaining sites")
nrow(records)


head(records)



 #filter site down to only sites that are present in records
sites <- sites %>%
  filter(site %in% unique(records$site))


print("number of sites with records data after filtering records data")
nrow(records)
 
```
## add columns to records to identify survey events and site-year combinations


```{r}

#create columns to identify survey events and site-year combinations
records<- records %>%
  mutate(survey.info = paste(site, year, month, day, week.no, sep = "_"),
         site.year.month = paste(site, year, month, sep = "_"),
         site.year = paste(site, year, sep = "_"))

```




## create a dataframe with one row for every year-site combination number of months for each site year (only when surveys carried out,no zeros)


```{r}
#find out how many months each site was surveyed each year and plot versus latitude
sum.months <- records %>% 
  distinct(site.year.month, .keep_all = TRUE) %>%
  group_by(site, year)%>%
  summarise(months.surveyed = n())

sum.months <- sum.months %>% 
      merge(
        select(
          sites, site, easting, northing), by = "site")  %>% 
  mutate(site.year = paste(site, year, sep = "_"))

table(as.character(sum.months$months.surveyed))

#small tendency for more northerly sites to be surveyed in fewer months
plot(months.surveyed~northing, data = sum.months)

#this is exploration, should be averaged per site first really
m1 <- lm(months.surveyed~northing, data = sum.months)

summary(m1)
```
## check distribution of number of surveys per site per year, plot number of sites surveyed per year in dataset, add years of surveys per site to site info
```{r}


#surveys per site per year
sum1 <-   records %>% 
  distinct(survey.info, .keep_all = TRUE) %>%
  group_by (site, year)  %>% 
summarise(n = n())

#inspect frequency distribution of surveys per site per year
print("surveys per site per year")
summary(sum1$n)
hist(sum1$n, main = "Surveys per site per year")

sum2 <- sum1 %>%
  mutate(n = as.character(n)) %>%
  group_by(n) %>%
      summarise(freq.survey = n())
sum2

#inspecting this shows that this site was surveyed multiple times per week in these years
check<- records %>%
  filter(site == 4831, year == 2019)%>% 
  distinct(survey.info, .keep_all = TRUE)

check<- records %>%
  filter(site == 4, year == 1976)%>% 
  distinct(survey.info, .keep_all = TRUE)

#get plot of sites surveyed per year by year
sites.per.year <- sum1 %>% 
  group_by(year)%>% 
summarise(sites = n())
print("sites per year")
summary(sites.per.year$sites)
plot(sites~year, data = sites.per.year)

#years per site summary and histogram
years.per.site <- sum1 %>% 
  group_by(site)%>% 
summarise(years.surveyed = n(),
          first.year = min(year),
          last.year = max(year))


print("years per site")
summary(years.per.site$years.surveyed)
hist(years.per.site$years.surveyed, xlim  = c(0, 50))


print("number of sites before merging operation")
nrow(sites)
sites <- merge(sites, years.per.site)  
print("number of sites after merging operation")
nrow(sites)
```

## Get vectors of sites time periods

```{r}

# a function to get a vector of site names for sites sampled in a r period
#x should be the above site.and.year
#condit is a condition to filter by


#filter_() might stop working, deprecated...
get.sites.in.date.range <- function(x, condit){
  sites.out <- x %>%
  filter_(condit)%>%
   dplyr::select(site)%>%
      distinct

sites.out <- sites.out[['site']]
  return(sites.out)
}

# dataframe of sites and years sampled
site.and.year <- dplyr::select(sum1, site, year)

sites.5.years <- get.sites.in.date.range(site.and.year, "year >= 2015 & year <=2019")


sites.1975 <- get.sites.in.date.range(site.and.year, "year == 1975")
sites.1980 <- get.sites.in.date.range(site.and.year, "year == 1980")
sites.1985 <- get.sites.in.date.range(site.and.year, "year == 1985")
sites.1990 <- get.sites.in.date.range(site.and.year, "year == 1990")
sites.1995 <- get.sites.in.date.range(site.and.year, "year == 1995")
sites.2000 <- get.sites.in.date.range(site.and.year, "year == 2000")
sites.2005 <- get.sites.in.date.range(site.and.year, "year == 2005")
sites.2010 <- get.sites.in.date.range(site.and.year, "year == 2010")
sites.2015 <- get.sites.in.date.range(site.and.year, "year == 2015")
sites.2016 <- get.sites.in.date.range(site.and.year, "year == 2016")
sites.2017 <- get.sites.in.date.range(site.and.year, "year == 2017")
sites.2018 <- get.sites.in.date.range(site.and.year, "year == 2018")
sites.2019 <- get.sites.in.date.range(site.and.year, "year == 2019")
sites.2022 <- get.sites.in.date.range(site.and.year, "year == 2022")


site.snapshots.to.map <- list(
  sites.1975,
  sites.1980, 
  sites.1985, 
  sites.1990, 
  sites.1995, 
  sites.2000, 
  sites.2005,
  sites.2010,
  sites.2015,
  sites.2018, 
  sites.2019, 
  sites.2022)


```

## Get records per species per year, filtered to only site-year combinations that meet certain criteria.

### months with survey records: >= 5

### records.site.year gives the number of records for each species in each site in each year

```{r}

#site year combinations to keep   
site.years.keep <- sum.months %>%
  dplyr::select(site.year, months.surveyed) %>%
  filter(months.surveyed >= 5)


#filter records per site per year to the desired site-year combinations
records.site.year <-  records %>%
  group_by (site, year, species.no)  %>% 
summarise(annual.records = sum(count))   %>% 
  mutate(site.year = paste(site, year, sep = "_"))   %>%
     merge(site.years.keep, #here we filter by merging, keeping only desired site-year combinations
           by = "site.year",
           all.y = TRUE)


#how many years per site in the filtered data?
sum.5.year <- records.site.year %>% 
  dplyr::select(site.year, site, year) %>%
  filter(year %in% 2015:2019)%>%
  distinct() %>%
  group_by(site) %>%
  summarise(years.15.19 = n())
  

table(sum.5.year$years.15.19)

print("number of sites before merging operation")
nrow(sites)
sites <- merge(sites, sum.5.year, by = "site", all.x = TRUE)
print("number of sites after merging operation")
nrow(sites)

```
```{r}
head(records.site.year)
```

# 2015-2019 data, combining T. sylvestris and T. lineola

```{r}

#to get the total records for each species in each site over  years:
records.site.5year <- records.site.year %>%
  filter(year %in% 2015:2019) %>% 
  group_by(site, species.no) %>%
  summarise(total.records = sum(annual.records))

records.site.5year$species.no <- paste("species", records.site.5year$species.no , sep = "_")

#get a site by species dataframe with counts of observations
site.by.obs.5.year.counts <- records.site.5year %>%
  pivot_wider(names_from = species.no,
              values_from = total.records) %>%
  column_to_rownames(var = "site")

# change NAs to zeros: these were cases when there wasn't a zero count record for a species that wasn't recorded in a survey
site.by.obs.5.year.counts[is.na(site.by.obs.5.year.counts)] <- 0

#aggregate Thymelicus lineola and T.sylvestris records

"#due to difficulties in field identification, we aggregated Essex Skipper Thymelicus lineola and Small #Skipper Thymelicus sylvestris into one taxon) "

site.by.obs.5.year.counts <- site.by.obs.5.year.counts %>%
  mutate(species_1000 = species_119 + species_120) %>%
  dplyr::select(-c(species_119, species_120))

#check if any sites have zero observations
any(rowSums(site.by.obs.5.year.counts) == 0)

#this was needed under previous filtering scheme, keepining in here in case needed again
site.by.obs.5.year.counts <- site.by.obs.5.year.counts[rowSums(site.by.obs.5.year.counts) > 0, ]

#make a presence absence site by species matrix
site.by.obs.5.year.presabs <- site.by.obs.5.year.counts

site.by.obs.5.year.presabs[site.by.obs.5.year.presabs[, ] > 0] <- 1
```

## make a row for combined species in species data 
```{r}

str(species)
#contents for new row
new.row <- c ("1000", "Combined Skippers", "Thymelicus lineola-sylvestris", "Wider countryside sp", "Grassland", 1)

print("number of species before merging two skippers")
nrow(species)
#add new row and remove two old rows
species <- rbind(species, new.row) %>%
  filter(species.no != "119" & species.no !="120")

print("number of species after merging two skippers")
nrow(species)
```



#check coverage

#sites filtered down to sites in 2015-2019
```{r}

#Transpose sity by counts matrix for use  by iNEXT package functions
check.coverage <- t(site.by.obs.5.year.counts)


#DataInfo() performs the first part of what iNEXT() does in seconds. the whole iNEXT() function would have to be left overnight for the whole dataset
#https://cran.r-project.org/web/packages/iNEXT/vignettes/Introduction.pdf

coverage <- DataInfo(check.coverage,  datatype="abundance")

hist(coverage$SC)

hist(coverage$SC, ylim = c(0, 50))


hist(coverage$SC, xlim = c(0.95, 1), breaks = 1000)

quantile(coverage$SC, seq(from = 0, to = 1, by = 0.01))


#it seems reasonable to include sites with greater than 0.95 coverage and surveyed over the flowering period

cov.merge <- coverage %>%
  dplyr::select(Assemblage, n, S.obs, SC)

names(cov.merge) <- c("site", "total.butterflies", "richness", "coverage")



print("number sites before filtering for time period")
nrow(sites)
sites.15.19 <- merge(sites, cov.merge, all.y = T)


print("number sites after filtering for time period")
nrow(sites.15.19)

```



#prepare additional site info from Rob, check which monitoring site data is present (UKBMS, NA)
```{r}
sites.more.info <- sites.more.info %>%
  dplyr::select(-c("Site Name", "Gridreference", "Easting", "Northing","No. yrs surveyed", "First year surveyed", "Last year surveyed"))
 
names(sites.more.info) <- c("site", "length", "country", "no.sections",  "survey.type" )



#66 sites  were not in site data from Rob , they must have been filtered out from the data he used for some reason
print("number of sites before merging operation")
nrow(sites.15.19)

sites.15.19 <- merge(sites.15.19, sites.more.info, by = "site", all.x = TRUE)

print("number of sites after merging operation")
nrow(sites.15.19)

sum3 <- sites.15.19 %>%
  group_by(survey.type) %>%
  summarise(n = n(),
            richness = mean(richness),
            butterflies = mean(total.butterflies),
            coverage = mean(coverage)
      )
sum3

```



```{r}
plot(richness~coverage, data = sites.15.19)


plot(coverage~years.15.19, data = sites.15.19)

```

## Plot Sites
```{r}

# plot.points <- st_as_sf(sites.15.19, coords=c("easting","northing"), remove = FALSE, crs = 27700)#I want to keep coords in the dataframe
# 
# #get a set of points for each unique grid cell
# 
# print("sites")
# nrow(plot.points)
# 
# basemap <-tm_shape(bounds) +
#   tm_borders()
# pointsmap <- basemap + 
# tm_shape(plot.points) +
#   tm_symbols(size = 0.05, col = "blue")
# 
# pointsmap2 <- basemap + 
# tm_shape(plot.points) +
#   tm_symbols(size = 0.5, col = "blue")
# 
# 
# print(pointsmap)
# print(pointsmap2)
```
```{r}
table(sites.15.19$country)

print("number of sites before filtering out channel islands, isles of man, lundy and scilly")
nrow(sites.15.19)

sites.15.19<- filter(sites.15.19,
               country != "Channel Isles" & country != "Isle of Man",
               site != 1791 & site != 1837)


print("number of sites after filtering out channel islands and isle of man, lundy and scilly")
nrow(sites.15.19)

table(sites.15.19$country)


#sites 1791 on Lundy in the Solent and 1837 on Scilly isles


```
## Plot Sites
```{r}
# 
# plot.points <- st_as_sf(sites.15.19, coords=c("easting","northing"), remove = FALSE, crs = 27700)#I want to keep coords in the dataframe
# 
# #get a set of points for each unique grid cell
# 
# print("sites.15.19")
# nrow(plot.points)
# 
# basemap <-tm_shape(bounds.2) +
#   tm_borders()
# pointsmap <- basemap + 
# tm_shape(plot.points) +
#   tm_symbols(size = 0.05, col = "blue")
# 
# pointsmap2 <- basemap + 
# tm_shape(plot.points) +
#   tm_symbols(size = 0.5, col = "blue")
# 
# 
# print(pointsmap)
# print(pointsmap2)
```


#filter to sites with sufficient coverage and effort

#filter site by species matrix to selected sites
```{r}

print("number of sites before filtering out adequate coverage and years of sampling")
nrow(sites.15.19)

sites.spatial <- sites.15.19 %>% 
                filter(coverage >= 0.9,
                years.15.19 >= 3)

print("number of sites after filtering out adequate coverage and years of sampling")
nrow(sites.spatial)

#this is the one to export!
site.by.obs.5.year.presabs1 <- site.by.obs.5.year.presabs %>%
  filter(rownames(site.by.obs.5.year.presabs) %in% sites.spatial$site)
```


## Plot Sites
```{r}

plot.points <- st_as_sf(sites.spatial, coords=c("easting","northing"), remove = FALSE, crs = 27700)#I want to keep coords in the dataframe
# 
# #get a set of points for each unique grid cell
# 
print("sites.15.19")
nrow(plot.points)

basemap <-tm_shape(bounds.2) +
  tm_borders()
pointsmap <- basemap +
tm_shape(plot.points) +
  tm_symbols(size = 0.05, col = "blue")
# 
# pointsmap2 <- basemap + 
# tm_shape(plot.points) +
#   tm_symbols(size = 0.5, col = "blue")
# 
# 
print(pointsmap)
# print(pointsmap2)
```



### get sites and years surveyed in more than 5 months 
```{r}

#years per site summary and histogram
years.per.site <- sum1 %>% 
  mutate(site.year = paste(site, year, sep = "_")) %>%
  filter(site.year %in% site.years.keep$site.year)%>%
  group_by(site)%>% 
summarise(years.surveyed.well = n(),
          first.year.well = min(year),
          last.year.well = max(year)) %>%
  ungroup() 

print("number of rows before merging")
nrow(sites)
sites.multiyear <- sites %>%
  merge(years.per.site, by = "site")
print("number of rows after merging")
nrow(sites)

str(years.per.site)

str(sites.multiyear)

```


```{r}
#get year by species matrix for all sites with greater than 40 years of records


#get a vector of the chosen sites
use.sites <- sites.multiyear %>%
  filter(years.surveyed.well >= 40)  %>%
    dplyr::select(site) %>%
     unlist() %>%
        as.numeric

#make a list to year by species matrices (which are actually dataframes)
year.by.species.matrices <- list()

for (i in 1:length(use.sites)){
   one.site.over.years.data <- records.site.year %>%
      filter(site == use.sites[i]) 
     
   
one.site.over.years.data$species.no <- paste("species", one.site.over.years.data$species.no, sep = "_")

   
    year.by.species.matrices[[i]] <-  one.site.over.years.data %>%
  
  pivot_wider(names_from = species.no,
              values_from = annual.records) %>%
  ungroup() %>%
  
  dplyr::select(-c(site, site.year, months.surveyed)) %>%
  
  column_to_rownames(var = "year")
    
    
 
#aggregate Thymelicus lineola and T.sylvestris records if both are present  
if("species_119" %in% names(year.by.species.matrices[[i]]) & "species_120" %in% names(year.by.species.matrices[[i]])){
year.by.species.matrices[[i]] <- year.by.species.matrices[[i]] %>%
  mutate(species_1000 = species_119 + species_120) %>%
  dplyr::select(-c(species_119, species_120))
}     
  }

head(year.by.species.matrices[[20]])
head(year.by.species.matrices[[43]])

dim(year.by.species.matrices[[43]])

#name the list elements after the sites
namevec <- vector()

length(namevec)<- length(use.sites)

  for (i in 1:length(use.sites)){
namevec[i] <- paste("site",  use.sites[i], sep = "_")
  }

names(year.by.species.matrices) <- namevec

#inspect output
str(year.by.species.matrices, max.level = 1)

#get site data frame for time series sites
sites.multiyear <- sites.multiyear %>%
  filter(site %in% use.sites)


check<- year.by.species.matrices[[1]]

```


## Plot Sites
```{r}

plot.points <- st_as_sf(sites.multiyear, coords=c("easting","northing"), remove = FALSE, crs = 27700)#I want to keep coords in the dataframe

#get a set of points for each unique grid cell

print("sites.15.19")
nrow(plot.points)

 basemap <-tm_shape(bounds.2) +
   tm_borders()
# pointsmap <- basemap +
# tm_shape(plot.points) +
#   tm_symbols(size = 0.05, col = "blue")

pointsmap2 <- basemap +
tm_shape(plot.points) +
  tm_symbols(size = 0.5, col = "blue")


# print(pointsmap)
print(pointsmap2)
```

# save processed data
```{r}


write_rds(site.year.by.species, file = here("./Data/Processed_Data/site_year_by_species.rds"))
write_rds(records, file = here("./Data/Processed_Data/processed_records.rds"))
write_rds(sites, file = here("./Data/Processed_Data/processed_sites.rds"))
write_rds(species, file = here("./Data/Processed_Data/processed_species.rds"))
write_rds(site.snapshots.to.map, file = here("./Data/Processed_Data/site_snapshots_to_map.rds"))
write_rds(use.sites, file = here("./Data/Processed_Data/sites_with_at_least_40_years.rds"))
write_rds(site.by.species.2015.to.2019, file = here("./Data/Processed_Data/site_by_species_2015_to_2019.rds"))

write_rds(site.by.species.2018, file = here("./Data/Processed_Data/site_by_species_2018.rds"))
write_rds(year.by.species.matrices, file = here("./Data/Processed_Data/year_by_species_matrices.rds"))
write_rds(surveys.site.year, file = here("./Data/Processed_Data/surveys_site_year.rds"))
```




