---
title: "Zeta Msgdm Results Processing"
output: html_document

params:
  dataset: "all_species"
  normalize_msgdm: "Simpson"
  orders_msgdm: 
   - 2
   - 4
   - 20
  seed: 4
---

```{r}
here::i_am("Scripts/Rmd_Scripts/process_msgdm_results.Rmd")

library(rmarkdown)
library(tidyverse)
library(zetadiv)
library(knitr)
library(knitrProgressBar)
library(tictoc)
library(gridExtra)
library(ggpubr)
library(mgcViz)
library(conflicted)
library(here)

conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::rename)
```

#### an example paper with helpful code for generating figures
```{r}
#https://onlinelibrary.wiley.com/doi/10.1111/mec.16860

# For forest plots, in future consider metator::forest, looks easier!!!
#https://www.metafor-project.org/doku.php/plots:forest_plot_with_subgroups
#https://bookdown.org/MathiasHarrer/Doing_Meta_Analysis_in_R/forest.html
```

#### File Paths
```{r}
# Path to data used for tje msgdm analysis (which is already done)
data_path <- paste0("./Data/Processed_Data/Spatial/", params$dataset, ".rds")

# Path to output of the msgdm analysis
msgdm_path <-    paste0("./Output/Spatial/msgdms", "/", params$dataset, "_" , 
                        params$normalize_msgdm,  "_" , params$seed, "/", 
                        params$dataset, "_" , params$normalize_msgdm, "_zeta_msgdms.rds")

# Path to folder to save output of this script
out_path <- here(paste0("./Output/Spatial/msgdms", "/", params$dataset, "_" , params$normalize_msgdm,  "_" , params$seed, "/", 
                       "/figs_tables//"))

if(!dir.exists(out_path)) dir.create(out_path, recursive = TRUE)
```

#### Load Data
```{r}
my_data <- read_rds(here(data_path))

my_msgdms <- read_rds(here(msgdm_path))

my_env_data <- my_data$site.by.env

# A vector of the orders for msgdms included
my_orders <- params$orders_msgdm
```

#### Extract processed ispline and environmental variable data into a dataframe.
```{r}
# Extract ispline data and environmantal data for each zeta order into a list
all_isplines <- list()
for(i in 1:length(my_orders)){
  temp_out <- Return.ispline(msgdm = my_msgdms[[i]], data.env = my_env_data, 
                  distance = TRUE) # Extract isplines and associated data
  # Normalised environmental data
  temp_env <- temp_out$env %>%
    rownames_to_column(var = "site")
  # Ispline predicted change in zeta diversity with environmental variable
  temp_isplines <- temp_out$Ispline
    names(temp_isplines) <- paste0("f.of.", names(temp_isplines)) #rename columns
    temp_isplines <- temp_isplines %>%
    mutate(zeta_order = paste0("zeta_", my_orders[i])) %>% #adding zeta_order as a column
    mutate(zeta_order = factor(zeta_order, levels = paste0("zeta_", my_orders))) %>%   
    relocate(zeta_order) %>%  
    cbind(temp_env) %>%
    relocate(site, .after = zeta_order) 
       
  all_isplines[[i]] <- temp_isplines
}


# Change from list to single dataframe, so that I can group by predictors for plots
all_isplines <- bind_rows(all_isplines) 
```

#### I have not yet managed to get plots of isplines against environmental data 
#### on the original scale
```{r}
# to do this, check downloaded code. If that doesn't help, add to questions for Guillaume


# head(temp_out$env)
# head(temp_out$env.num)
# 
# head(temp_env)
# 
# head(temp_isplines)


# mins <- temp_out$env.num %>%
#   summarise(across(where(is.numeric), min))
# maxs <- temp_out$env.num %>%
#   summarise(across(where(is.numeric), max)) 
# 
# env_diffs <- data.frame(t(maxs-mins))
# names(env_diffs) <- "max_possible_difference"

```

#### Plot isplines for each predictor for different zeta orders
```{r}
for(i in 1:(ncol(my_env_data) + 1)){
  print(
    ggplot(data = all_isplines, aes(x = all_isplines[, (i + 11)], 
                                  y = all_isplines[,(i + 2)], group = zeta_order)) +
    geom_line(aes(colour = zeta_order))  + 
    xlab(names(all_isplines)[i + 11]) + 
    ylab("Change in Zeta diversity")
  )
}
```

### get data to to make multipanel plot for BES using only zeta = 4
```{r}

#subset to zeta = 4 and get maximum y axis for value for ispline plots
subs_isplines <- all_isplines %>%
  filter(zeta_order == "zeta_4") %>%
  rename(Temperature = mean.ann.temp,
         Log_Prec. = log10.mean.ann.rain,
         Top_Wet. = topographic.wetness,
         pH = ph,
         Nit_Dep. = total.n,
         Tree_Dens. = tree.density,
         Log_Pestic. = log10.pesticide.risk,
         Log_Humans = log10.humans,
         Geog_Dist = distance
  )

names(subs_isplines)

check <- subs_isplines[,3:11]

# make the y axis a little taller than largest value over all variables

#I have decided to leave distance off the plot for BES
max_y <- 1.05*max(subs_isplines[, 3:10])
```

#### Make ispline plots for each variable and store in a list
```{r}

# see explanation for code to store ggplot objects in list here: https://stackoverflow.com/questions/31993704/storing-ggplot-objects-in-a-list-from-within-loop-in-r

my_plots <- vector('list', (ncol(my_env_data) + 1))
names(my_plots) <- c("p1","p2" ,"p3", "p4","p5" ,"p6" ,"p7" ,"p8" ,"p9")

for (i in 1:(ncol(my_env_data) + 1)) {
    message(i)
    my_plots[[i]] <- local({
      i <- i
      plot_i <- ggplot(data = subs_isplines, aes(x = subs_isplines[, (i + 11)], 
                                  y = subs_isplines[,(i + 2)])) +
      geom_line(colour = "red") +
  ylab("") +
  xlab(names(subs_isplines)[i + 11]) + 
  theme_classic()  +
  theme(axis.title = element_text(size = 5),
        axis.text = element_text(size = 4),
  #      axis.text.y = element_blank()
  ) + 
  coord_cartesian(ylim = c(0, max_y)) +
  theme(plot.title = element_text(size = 5.5))
      print(plot_i)

   
    })
}



```
# Make a multipanel plot
```{r}
print(c("max y = ", max_y))
# Export plots from the list to the global environment
list2env(my_plots, envir = .GlobalEnv)

# mgcViz function for multipanel plots
gridPrint(p1, p2, p3, p4, p5, p6, p7, p8,
          respect=TRUE,
          ncol = 8)

# as above, write to file
png(file = paste0(out_path, "msgdm_multiplot1.png"), pointsize=10, width=9000, height=1000, res=600)
gridPrint(p1, p2, p3, p4, p5, p6, p7, p8,
          respect=TRUE,
          ncol = 8)
dev.off()


```
#### Extract coefficients for each zeta order
```{r}
# Extract coefficients for each zeta order into a list
all_coef <- list()
for(i in 1:length(my_orders)){
  temp_summary <- summary(my_msgdms[[i]]$model)
  
  # info needed to assign values below
  coef_rows <- nrow(temp_summary$coefficients)
  number_of_colours <- 1 + (nrow(temp_summary$coefficients) -1 )/3
  
  temp_coef <- rep(paste0("zeta_order", i), nrow(temp_summary$coefficients)) %>%
    data.frame() %>%
    rename(zeta_order = ".") %>%
    cbind(temp_summary$coefficients) %>%
    rename(p_value = "Pr(>|z|)",
      z_value = "z value",
      Std_Error = "Std. Error") %>% 
    rownames_to_column(var = "Effect")   %>% 
    mutate(Effect = factor(Effect, levels = Effect),
    zeta_order = paste0("zeta_", my_orders[i]),#adding zeta_order as a column
      Lower = Estimate - 1.96*Std_Error, #CL
      Upper = Estimate + 1.96*Std_Error,
      colours = factor(c(1, rep(2:number_of_colours, each = 3))),
      y_height = as.numeric(rev(seq(1,coef_rows))))
  # tried to add a row so this aligns with forest plot, will change text colour to white later
  #blank <- temp_coef[1,]
  #blank$Effect <- "blank"
  #temp_coef <- rbind(blank, temp_coef)
  
  
  all_coef[[i]] <- temp_coef
}



check<- bind_rows(all_coef)
```

#### Make a forest plot for each zeta order and store in a list.
```{r}
forest_plots <- list()

for (i in 1:length(my_orders)){

plot_data <- all_coef[[i]] 

p <- ggplot(plot_data, aes(x = Estimate, y = y_height)) +
  geom_point(aes(colour = colours)) +                                     
  geom_pointrange (aes(xmin = Lower, xmax = Upper, colour = colours)) + 
  scale_y_continuous(name = NULL, breaks = plot_data$y_height,
    labels = plot_data$Effect) +
  ggtitle(paste0("Zeta = ", my_orders[i])) +
  theme_bw() +
  theme(panel.border = element_blank(),
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line.x = element_line(),
    axis.line.y = element_blank(),
    axis.text.x = element_text(colour = "black", size = 11),
    axis.text.y = element_blank(),
    #axis.text.y = element_text(colour = "black", size = 10),
    axis.ticks.x = element_line(),
    axis.ticks.y = element_blank(),
    axis.title.x = element_text(), 
    plot.title = element_text(hjust = 0, face = "bold"),
  legend.position="none")  +  
  geom_vline(xintercept = 0, color = "red",linetype = "dashed", alpha = 0.5) 

forest_plots[[i]] <- p
}

```

# Make a table with names of predictor splines and intercept to add to forest plot
```{r}
table_base <- 
  ggplot(data = all_coef[[1]], aes(y = y_height)) + # everything in this plot is empty 
  ylab(NULL) + xlab('') + ggtitle('') + 
  ylim(0, 28) + # make sure this is the same as p1
  theme(panel.border = element_blank(),
    panel.background = element_blank(),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.text.x = element_text(color = "white", hjust = -3, size = 11),
    axis.text.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0, face = "bold"),# make sure hjust is the same as p1
    legend.position = "none") 
```

#### A table for effect names (to be improved for presentation), font size 10
```{r}
effects_table <- table_base + 
  geom_text(aes(y = y_height, 
                x = 0, 
                colour = colours,
                label = Effect), 
            size = 3.1, hjust = 1, vjust = -0.5) +
            xlim(-2,0) +
            ggtitle("Effect")  + 
            theme(plot.title = element_text(size = 10, face = "bold"))
```


#### A table for effect names (to be improved for presentation), default font size
```{r}
effects_table2 <- table_base + 
  geom_text(aes(y = y_height, 
                x = 0, 
                colour = colours,
                label = Effect), 
            size = 3.1, hjust = 1, vjust = -0.55) +
            xlim(-2,0) +
            ggtitle("Effect")
```

#### A list of tables of p values to combine with forest plots
```{r}
p_tables <- list()

for (i in 1:length(my_orders)){
p_table <- table_base + 
  geom_text(data = all_coef[[i]],
    aes(y = y_height, 
        x = 0, 
        colour = colours,
        label = paste0(format(round(p_value, digits = 3), nsmall = 3))), 
        size = 3.1, hjust = 0, vjust = -0.5)  +
        xlim(0,1) +
        ggtitle('')

p_tables[[i]] <- p_table
}
```


#### A list of tables of p values with headings to combine with forest plots
```{r}

p_tables2 <- list()

for (i in 1:length(my_orders)){
p_table <- table_base + 
  geom_text(data = all_coef[[i]],
    aes(y = y_height, 
        x = 0, 
        colour = colours,
        label = paste0(format(round(p_value, digits = 3), nsmall = 3))), 
    size = 3.1, hjust = 0, vjust = -0.5)  +
    xlim(0,1) +
    ggtitle(paste0("Zeta = ", my_orders[i])) + 
    theme(plot.title = element_text(size = 10, face = "bold"))

p_tables2[[i]] <- p_table
}
```

#### Separate forest plots with p values for each zeta order
```{r}
for (i in 1:length(my_orders)){
grid.arrange(effects_table, p_tables[[i]], forest_plots[[i]], 
             nrow = 1, ncol = 3)
}
```

#### Makes lists for plotting forest plots for different isplines together
```{r}
# Labels for effects with the heading in different font sizes to match
effects_table <- list(effects_table)
effects_table2 <- list(effects_table2)

# List for forest plot
forest_plots <- c(effects_table2, forest_plots)

# List of p values corresponding to forest plots for plotting separately
p_tables <- c(effects_table, p_tables)
# List of p values corresponding to forest plots for plotting together
p_tables2 <- c(effects_table, p_tables2)
```

#### Forest plot
```{r}
grid.arrange(grobs = forest_plots,
              nrow = 1, ncol = length(forest_plots))

filepath <- paste0(out_path, "/forest_plot", ".png")
unlink(filepath, recursive = FALSE, force = FALSE) #deletes previous version with the same name
png(file = filepath, 
   width = 55*length(forest_plots),
   height = 210, 
   units = "mm", 
   res = 72)
grid.arrange(grobs = forest_plots,
              nrow = 1, ncol = length(forest_plots))

dev.off()
```

#### Table of p values for all zeta orders
```{r}
grid.arrange(grobs = p_tables2,
              nrow = 1, ncol = length(p_tables))

filepath <- paste0(out_path, "/forest_plot_p_values", ".png")
unlink(filepath, recursive = FALSE, force = FALSE) #deletes previous version with the same name
png(file = filepath, 
   width = 55*length(p_tables2),
   height = 210, 
   units = "mm", 
   res = 72)
grid.arrange(grobs = p_tables2,
              nrow = 1, ncol = length(p_tables2))
dev.off() 
```

#### Check if we can retrieve the same ispline plots
```{r}
my_msgdm_i <- my_msgdms[[1]]

my_data.env <- my_data$site.by.env

my_isplines <- Return.ispline(msgdm = my_msgdm_i, data.env = my_data.env, distance = TRUE)

# These two are almost identical except that the quantiles for distance are 
# slightly different
Plot.ispline(msgdm = my_msgdm_i, data.env = my_data.env, distance = TRUE)
Plot.ispline(isplines = my_isplines)

str(my_isplines, max.level = 1)
str(my_msgdms, max.level = 1)

isp_env <- my_isplines$env

head(my_isplines$Ispline)
```


#### Deviance explained (rough as this is a non-linear model)

```{r}
dev_table <- data.frame()

for (i in 1:length(params$orders_msgdm)) {
dev_table[i, 1] <- paste0("Zeta = ", my_orders[[i]])
dev_table[i, 2] <- with(summary(my_msgdms[[i]]$model), 1 - deviance/null.deviance)
}
names(dev_table) <- c("Zeta order", "Deviance Explained")

#I have checked and csv over-writes fines
write_csv(dev_table, file = paste0(out_path, "/", "deviance_table.csv"))
kable(dev_table, digits = 3)
```

#### Variation explained table (also rough)
```{r}
# make the table for output

# First column is the source of variation
source <- c("Environment", "Distance", "Environment or Distance", "Total")
varparts <- data.frame(source)
names(varparts) <- "Source of Variation"

# remaining columns are the variation explained for each zeta order
for (i in 1:length(params$orders_msgdm)) {
  set.seed(1)
  zetavar <- Zeta.varpart(msgdm.mod = my_msgdms[[i]], method.glm = "glm.fit2", cons.inter = -1)
  zetavar <- 100*zetavar[, 1]  %>%
  round(3)
#formatting step needed for digits after decimal
zetavar <- formatC(zetavar, 1, format = "f")
  
varparts[, i+1] <-  c( zetavar[6], zetavar[4], zetavar[5], zetavar[1])
names(varparts)[i + 1] <- paste0("Percentage Explained for Zeta order ", params$orders_msgdm[i])

}

write_csv(varparts, file = paste0(out_path, "variance_partitioning_table.csv"))

```


#### Below is code for validation



<!-- ```{r} -->
<!-- Zeta.order.mc -->
<!-- ``` -->


<!-- ## try to set up k-fold cross validation -->

<!-- ## first an msgdm for training data -->
<!-- ```{r} -->
<!-- set.seed(1) -->
<!-- sites$fold <- kfold(sites, k=5, by = sites$country) -->

<!-- #this shows that the "by"argument works as I understand. It stratifies the sampling -->
<!-- sum <- sites %>% -->
<!--   group_by(fold, country) %>% -->
<!--   summarise(n= n()) -->

<!-- #subset to first fold -->
<!-- S.SP <- site.by.species[sites$fold != 1 ,] -->

<!-- S.XY <- site.by.xy[sites$fold != 1 ,] -->

<!-- S.E <- site.by.env[sites$fold != 1 ,]   -->



<!-- set.seed(1) -->
<!-- msgdm.train <-  Zeta.msgdm(S.XY, data.spec = S.SP, data.env = S.E,  order=2, sam=5000, -->
<!--   reg.type="ispline", normalize="Simpson", family=binomial(link="log"),cons.inter = -1, glm.init = TRUE)  -->


<!-- ``` -->
<!-- ## here i do an entire msgdm for the holdout data because I can't see another way to get the transformed values for distance -->

<!-- ```{r} -->

<!-- S.SP1 <- site.by.species[sites$fold == 1 ,] -->

<!-- S.XY1 <- site.by.xy[sites$fold == 1 ,] -->

<!-- S.E1 <- site.by.env[sites$fold == 1 ,]   -->

<!-- z.test2 <- Zeta.msgdm(S.XY1, data.spec = S.SP1, data.env = S.E1,  order=2, sam=5000, -->
<!--   reg.type="ispline", normalize="Simpson", family=binomial(link="log"),cons.inter = -1, glm.init = TRUE) -->

<!-- test.predictors<- z.test2$predictors -->



<!-- pred.zeta <- Predict.msgdm(model.msgdm = msgdm.test$model, reg.type = "ispline", newdata =  test.predictors) -->

<!-- m1 <- lm (z.test2$val ~ pred.zeta) -->

<!-- summary(m1) -->
<!-- anova(m1) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- Plot.ispline(msgdm = msgdm.train, data.env = S.E, distance = TRUE,  legend = TRUE) -->

<!-- Plot.ispline(msgdm = z.test2, data.env = S.E1, distance = TRUE,  legend = TRUE) -->

<!-- ``` -->
<!-- ```{r} -->
<!-- #Zeta.varpart -->

<!-- print("ZETA 2") -->
<!-- summary(msgdm.train$model) -->
<!-- summary(z.test2$model) -->


<!-- ``` -->

<!-- ```{r} -->

<!-- summary(msgdm.train$model) -->
<!-- summary(z.test2$model) -->
<!-- print("variance explained zeta 2") -->
<!-- with(summary(msgdm.train$model), 1 - deviance/null.deviance) -->
<!-- with(summary(z.test2$model), 1 - deviance/null.deviance) -->
<!-- ``` -->








<!-- ```{r} -->
<!-- z2 <- Zeta.order.mc(data.spec = S.SP, xy = S.XY, sam = 2000, order = 2, normalize = "Simpson") -->
<!-- str(z2) -->
<!-- ``` -->

<!-- ## we need to get the Isplines, which ar -->


<!-- ## use ispline to transfrom data for input to predict -->



<!-- ```{r} -->

<!-- ``` -->



