---
title: "R Notebook"
output: html_notebook
---

Previous version of code in filter_timeseries.Rmd

This is abbreviated 

so far I have only used this to choose three sites for Cang to play with

#### Load packages.
```{r}
here::i_am("Scripts//Rmd_Scripts//process_temporal_data.Rmd")

library(conflicted)

library(tidyverse)

library(sgo)

library(sf)

library(tmap)

library(here)

library(iNEXT)

library(tictoc)

conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::mutate)
conflicts_prefer(dplyr::select)
conflicts_prefer(base::split)
```


#### Functions
```{r}
source(here("Scripts/R_scripts/butterfly_richness_turnover_functions.R"))

```

#### File Paths
```{r}
# Path to data used for tje msgdm analysis (which is already done)
data_path <- paste0("./Data/Processed_Data/Temporal/for_temporal_ukbms.rds")


# Path to folder to save output of this script
out_path <- here(paste0("./Data/Processed_Data/Temporal/"))

#if(!dir.exists(out_path)) dir.create(out_path, recursive = TRUE)
```

#### Load Data
```{r}
my_data <- read_rds(here(data_path))

records <- my_data$records

species <- my_data$species

sites <- my_data$sites

# Basemap without outlying islands
GB_simpl <- readRDS(here("./Data./Maps./gb_multipolygon_simplified.rds"))
```

## Plot Sites
```{r}
plot.points <- st_as_sf(sites, coords=c("x","y"), remove = FALSE, crs = 27700) 

basemap <-tm_shape(GB_simpl) +
  tm_borders()

pointsmap <- basemap +
tm_shape(plot.points) +
  tm_symbols(size = 0.05, col = "blue")

pointsmap2 <- basemap +
tm_shape(plot.points) +
  tm_symbols(size = 0.5, alpha = 0.5, col = "blue")

print(pointsmap)
print(pointsmap2)
```

#### Make a table to summarise remaining data and check data filtering at each step
```{r}
# Make a table
prisma <- tibble(
"Step" = "Loaded Data",
"Records" = nrow(records),
"Site.years in records" = length(unique(records$site.year)),
"Sites in records" = length(unique(records$site)),
"Species in records" = length(unique(records$species.no)),
"Sites in sites" = nrow(sites),
"Species in species" = nrow(species)
)

prisma
```


# filter by by time to pre-covid data
```{r}
records <- records %>%
  filter(year <= 2019)

filtered_data <- filter_ukbms("records")
# This updates the objects in the global environment with filtered versions
list2env(filtered_data, envir = .GlobalEnv)

prisma <- add_prisma(step_name = "Filtered to records before covid")
prisma
```

#### Filter to site-year combinations sampled in all four sampling periods
```{r}
# Get a dataframe with the number of surveys per period in each site in each year
surveys.by.period <- records %>%
  distinct(survey.info, .keep_all = TRUE) %>%
  group_by(site, year, sampling.period) %>%
  summarise(surveys.per.period = n()) %>%
  pivot_wider(names_from = sampling.period,
    values_from = surveys.per.period,
    values_fill = 0) %>%
  mutate(site.year = paste("site", site, year, sep = "_"), 
         #criterion for inclusion, being sampled in all 4 periods
         all.4 = per1 > 0 & per2 > 0 & per3 > 0 & per4 > 0,
         total.surveys = sum(per1, per2, per3, per4)) %>% 
  ungroup()

head(surveys.by.period)
```

```{r}
# Get number of years for each site that there were surveys in all four periods
sites.years.with.8plus.surveys  <- surveys.by.period %>% 
  filter(total.surveys >= 8) 

# this merge adds info on number of years with all four periods and filters
# to only include years with two or more sites
records <- records %>%
  filter(site.year %in% sites.years.with.8plus.surveys$site.year)

print("filter to site.years with at least 8 surveys")
filtered_data <- filter_ukbms("sites")
list2env(filtered_data, envir = .GlobalEnv)

prisma <- add_prisma(step_name = "Filtered to site.years with at least 8 surveys")
tail(prisma)
```


```{r}
# Get number of years for each site that there were surveys in all four periods
sites.years.with.all.four  <- surveys.by.period %>% 
  filter(all.4 == TRUE) 

# this merge adds info on number of years with all four periods and filters
# to only include years with two or more sites
records <- records %>%
  filter(site.year %in% sites.years.with.all.four$site.year)

print("filter to site.years sampled in all four periods")
filtered_data <- filter_ukbms("sites")
list2env(filtered_data, envir = .GlobalEnv)

prisma <- add_prisma(step_name = "Filtered all 4 periods")
tail(prisma)
```



#### get surveys for each site in each year in each of sampling periods that I define
```{r}
#surveys per site per year
surveys.site.year <-   records %>% 
  distinct(survey.info, .keep_all = TRUE) %>%
  group_by (site, year, sampling.period)  %>% 
  summarise(n = n())  %>% 
  ungroup() %>%
  mutate(site.year = paste("site", site, year, sep = "_")) %>%
  pivot_wider(
    names_from = sampling.period,
    values_from = n,
    values_fill = 0) %>%
  mutate(surveys.per.year = (per1 + per2 + per3 + per4))

head(surveys.site.year)

#inspect frequency distribution of surveys per site per year
print("surveys per site per year")
summary(surveys.site.year$surveys.per.year)
hist(surveys.site.year$surveys.per.year, main = "Surveys per site per year")



```


#### Obtain dataframe with number of years per site, and minimum surveys for each site per year and  in each sampling period in each year
```{r}
#years per site summary and histogram
years.per.site <- surveys.site.year %>% 
  group_by(site)%>% 
summarise(years.surveyed = n(),
          first.year = min(year),
          last.year = max(year),
          span.surveyed = last.year - first.year,
          missing.years = span.surveyed + 1 - years.surveyed,
          minimum.surveys.per.year = min(surveys.per.year))

head(years.per.site)

#inspect frequency distribution of years
print("years per site")
summary(years.per.site$years.surveyed)
hist(years.per.site$years.surveyed, main = "Years per site")
```

#### Obtain site by species (count) matrix and calculated coverage
```{r}

# This is done with a function as there is a large amount of code that must be run again after filtering. 
# Argument is the records dataframes, 
# Steps are 1) aggregate to get count of each species in each site-year
#           2) get the site by species dataframe for counts
#           3) calculate coverage for each site year
#           4) calculates a variable for time since last survey in each site
#           5) summarises time series for each site
# It prints the result of each of these steps and also checks that there are no sites or species without records
# It outputs results of steps 2-5 in list

#function to calculate the variable "diff" in function below
get_diff <- function(df){
  diff_vec <- diff(df$year)
  df$diff <- c(0, diff_vec)
  return(df)
}

# get dataframe with records per species for each site-year combination
get_site_by_species_and_coverage <- function (my_records){
  aggregated.records <- my_records %>%
  group_by(site, year, species.no)%>%
  summarise(annual.records = sum(count))
print(head(aggregated.records))   
  
#get a site.year by species dataframe with counts of observations
site.year.counts <- aggregated.records %>%
  pivot_wider(names_from = species.no,
              values_from = annual.records,
              values_fill = 0) %>% # adds zero counts back in (these were filtered earlier)
  mutate (site.year = paste("site", site, year, sep = "_")) %>% 
  column_to_rownames(var = "site.year")%>%
  select(-c(site, year))
names(site.year.counts) <- paste("species", names(site.year.counts), sep = "_")
print(head(site.year.counts))

print("are there any species with no records")
print(any(colSums(site.year.counts) == 0))

print("are there any sites with no records")
print(any(rowSums(site.year.counts) == 0))

#calculate coverage
coverage.multi <- DataInfo(t(site.year.counts),  datatype="abundance")
coverage.multi <- coverage.multi %>%
  mutate(site = str_split_i(Assemblage, "_", 2),
         site = paste0("site_", site),
         year = str_split_i(Assemblage, "_", -1),
         .before = Assemblage) %>% 
  rename(site.year = Assemblage)

print(head(coverage.multi))


# get data for each site.year on years since last data in dataset for that site
site.year.cov <- coverage.multi[, 1:6]
# split by site.year into a list of files
split_data <- site.year.cov %>%
  mutate(year = as.integer(year), 
         site = as.factor(site)) 
split_data <- base::split(split_data, split_data$site)

# Calculate the variable "diff"
split_data <- lapply(split_data, get_diff)
# combine the list back into a single dataframe
site.year.cov  <- bind_rows(split_data)
print(head(site.year.cov))

#summarise time series by site
site.sum <- site.year.cov %>%
  group_by(site) %>% 
  summarise(
    total.years = n(),
    max.gap = max(diff) - 1,
    missing.years = sum(diff) - total.years +1)
print(head(site.sum))

data_out <- list(site.year.counts, coverage.multi, site.year.cov, site.sum)
names(data_out) <- c("site.year.counts", "coverage.multi", "site.year.cov", "site.sum")

return(data_out)
}
```

#### Run the function
```{r}
site.by.species.list <- get_site_by_species_and_coverage(records) 
```

#### Inspect coverage
```{r}
coverage.multi <- site.by.species.list[["coverage.multi"]]

hist(coverage.multi$SC)

hist(coverage.multi$SC, xlim = c(0.98, 1), breaks = 200)

hist(coverage.multi$SC, xlim = c(0.88, 0.99), ylim = c(0, 1000), breaks = 150)

quantile(coverage.multi$SC, seq(from = 0, to = 1, by = 0.01), na.rm = TRUE)

summary(coverage.multi$SC)
```

# filter to site years with at least 0.99 coverage
```{r}
site.year.cov <- site.by.species.list[["site.year.cov"]]

# get a dataframe with the site-year combinations with >=99 % coverage
site.year.cov.filtered <- site.year.cov %>%
  filter(SC >= 0.99)

records <- filter(records, site.year %in% site.year.cov.filtered$site.year)

filtered_data <- filter_ukbms("records")
# This updates the objects in the global environment with filtered versions
list2env(filtered_data, envir = .GlobalEnv)

prisma <- add_prisma(step_name = "Filtered to site.years with at least 0.99 coverage")
prisma

head(records)
```

#### Filter to sites with >= 20 years data (after meeting othe filtering criteria)
```{r}
enough.years <- filter(years.per.site, years.surveyed >= 20)

records <- filter(records, site %in% enough.years$site)

filtered_data <- filter_ukbms("records")
# This updates the objects in the global environment with filtered versions
list2env(filtered_data, envir = .GlobalEnv)

prisma <- add_prisma(step_name = "Filtered to sites with >= 20 years")
prisma
```

# rerun function to get site by species and coverage 
```{r}
site.by.species.list <- get_site_by_species_and_coverage(records) 
```



# redo coverage matrix

```{r}
site.year.cov <- site.by.species.list[["site.year.cov"]]

cov_mat <- site.year.cov %>%
  select(site, year, SC) %>%
  pivot_wider(values_from = SC, names_from = site) %>%
  column_to_rownames(var = "year") %>%
  as.matrix()

dim(cov_mat)

heatmap(cov_mat[, 1:55], Rowv= NA, Colv = NA)
heatmap(cov_mat[, 56:110], Rowv= NA, Colv = NA)
heatmap(cov_mat[, 111:165], Rowv= NA, Colv = NA)
heatmap(cov_mat[, 166:221], Rowv= NA, Colv = NA)
```



## Plot Sites
```{r}
plot.points <- st_as_sf(sites, coords=c("x","y"), remove = FALSE, crs = 27700) 

basemap <-tm_shape(GB_simpl) +
  tm_borders()

pointsmap <- basemap +
tm_shape(plot.points) +
  tm_symbols(size = 0.05, col = "blue")

pointsmap2 <- basemap +
tm_shape(plot.points) +
  tm_symbols(size = 0.5, alpha = 0.5, col = "blue")

print(pointsmap)
print(pointsmap2)
```

```{r}
site.year.by.species.counts <- site.by.species.list[["site.year.counts"]]

#make a presence absence site by species matrix
site.year.by.species.presabs <- site.year.by.species.counts
site.year.by.species.presabs[site.year.by.species.presabs[, ] > 0] <- 1

head(site.year.by.species.counts)
head(site.year.by.species.presabs)
```

#### Extract data on number of sites in 2015-2019 data before and after subsequent filtering
```{r}
#prisma_short <- prisma[c(9, 12, 13), c(1,3) ]

#prisma_short$Step <- c("All records", "Adequate effort", "Adequate coverage")

#names(prisma_short) <- c("Records", "Sites Remaining")

#prisma_short$`Sites Dropped` <- c(NA, abs(diff(prisma_short$`Sites Remaining`)))

```

```{r}
temporal_ukbms_obs_sites_spp <- list(site.year.by.species.counts, site.year.by.species.presabs, species, sites)
names(temporal_ukbms_obs_sites_spp) <- c("site.year.by.species.counts", "site.year.by.species.presabs",  "species", "sites")

write_rds(temporal_ukbms_obs_sites_spp, file = here("./Data/Processed_Data/Temporal/temporal_ukbms_obs_sites_spp.rds"))

csv_path <- paste0(here("Output/Temporal/filtering",  "/"))
if(!dir.exists(csv_path)) dir.create(csv_path)

write_csv(prisma, file = paste0(csv_path, "/", "filtering_summary.csv"))
#write_csv(prisma_short, file = paste0(csv_path, "/", "sites_filtering_summary.csv"))
```




<!-- #### Filter to include three sites as requested by Cang -->
<!-- ```{r} -->
<!-- records <- records %>% -->
<!--   filter(site %in% c(1, 4, 86)) -->

<!-- three_series <- filter_ukbms("records") -->

<!-- ``` -->


<!-- #### Check all sites and species have occurrence data -->
<!-- ```{r} -->
<!-- print("are there any species with no records") -->
<!-- any(colSums(site.year.counts)==0) -->

<!-- print("are there any sites with no records") -->
<!-- any(rowSums(site.year.counts)==0) -->

<!-- print("sites and species in matrix") -->
<!-- dim(site.year.counts) -->
<!-- ``` -->
<!-- #### move info out of column and row names into the dataframe to save as csv -->

<!-- ```{r} -->
<!-- site.year.counts <- site.year.counts %>% -->
<!--   rownames_to_column(var = "site_year") %>% -->
<!--   mutate(site = str_sub(site_year, end = -6), -->
<!--          year = str_sub(site_year, -4)) %>% -->
<!--   relocate(site, year, .after = site_year) -->
<!-- ``` -->



<!-- #### write csv to processed data file -->

<!-- ```{r} -->
<!-- write_csv(site.year.counts, -->
<!--           col_names = TRUE, -->
<!--           file = paste0(out_path, "/", "three_sites_site_year_counts.csv")) -->
<!-- ``` -->











