---
title: "Process_butterflies"
output: html_notebook
---



## Data cleaning, obtaining site by species matrices
```{r}

here::i_am("Scripts//Rmd_Scripts//Process_CSV_data.Rmd")

library(conflicted)
library(tidyverse)
library(here)

#library(foreach)
#library(doParallel)
#library(doSNOW)

library(spaa)

library(sgo)


library(iNEXT)

conflicts_prefer(snow::stopCluster)
conflicts_prefer(snow::makeCluster)
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::mutate)

#filter <- dplyr::filter


```


```{r}
#(cl = makeCluster(parallel::detectCores()-2, type = "SOCK")) # Initialize all cores except for two
#registerDoSNOW(cl)


records <- read_csv(here("./Data/CSV_Data/ukbms_Rogers_10-3-23.csv"))

#In the CSV, I  added in column headings, did a sort and added in extra columns to help correct entries in incorrect columns
#note, some sites have incomplete info, one site remains with latitude and longitude in wrong columns and no other info, such sites are filtered in this file


sites <- read_csv(here("./Data/CSV_Data/ukbms_Rogers_sites_10-3-23_to_correct.csv"))

sites.more.info <- read_csv(here("./Data/CSV_Data/ukbmsSiteLocationData1976-2019.csv"))

species <- read_csv(here("./Data/CSV_Data/ukbms_Rogers_species_10-3-23.csv"))

```

```{r}
head(records)
summary(records$YEAR)
```
## this shows that column names need attention

## inspecting the dataframe shows some info is in the wrong columns
```{r}
head(sites)
nrow(sites)
```

## clean and organise site data
```{r}

#rename columns
names(sites)
sites <- sites %>%
  rename(site = SITENO,
         site.name =  SITENAME,
         gridref = GRIDREF,
         longitude = LONGITUDE,
         latitude = LATITUDE) %>% 
  #here we move values that are in the wrong columns to the correct columns
  mutate(site.name = case_when(paste2cols == 1 ~ paste(site.name, gridref, sep = "_"),
                             paste3cols == 1 ~ paste(site.name, gridref, longitude, sep = "_"),
                             .default = as.character(site.name)),
         gridref = case_when(paste2cols == 1 ~ as.character(longitude),
                             paste3cols == 1 ~ as.character(latitude),
                             .default = as.character(gridref)),
         longitude = case_when(paste2cols == 1 ~ as.numeric(latitude),
                             paste3cols == 1 ~ as.numeric(extracol1),
                             .default = as.numeric(longitude)), 
         latitude = case_when(paste2cols == 1 ~ as.numeric(extracol1),
                             paste3cols == 1 ~ as.numeric(extracol2),
                             .default = as.numeric(latitude))) %>% 
  select(-c(extracol1, extracol2, paste2cols, paste3cols))


# for inspection only, sites that have no co-ordinates or grid square reference
bad.sites <- filter(sites, is.na(longitude) == T &  is.na(gridref) ==T)

#remove irish sites, which have uneven number of characters in the gridref (they start with a single letter)
sites <- sites %>%
                mutate(str.len.gridref = str_length(gridref),# a variable showing number of characters in gridref
                       even = str.len.gridref %% 2 == 0)  #whether above variable is even


irish.sites <- filter(sites, even == F)  #for inspection only

#can be used
sites <-  filter(sites, is.na(longitude) == F |  is.na(gridref) == F,#remove 51 sites with neither a gridref nor longlat
                even == T) %>% #filter 138 Irish sites
                    select(-c(even, str.len.gridref))

#one wonky site not picked up above
sites <- filter(sites, gridref != "Lurgan") 


         
```


```{r}
dplyr::last_dplyr_warnings()
head(species)
nrow(species)
```
#add in missing longitude and latitude for 20 sites
```{r}

#obtain longitude and latitude using sgo

#just work with rows with missing dtaa
for.longlat <- sites %>% 
  filter(is.na(longitude) == TRUE)

#get eastings and northings first
for.longlat <- sgo_ngr_bng(for.longlat, col = "gridref")

#get latitude and longitude
for.longlat <- sgo_bng_lonlat(for.longlat)

#get a dataframe of rows with missing dat to add them into
check <-sites %>%
filter(is.na(longitude) == TRUE) %>%
  rename(long = longitude, 
         lat = latitude)

#add in longitude and latitude
check$long <- for.longlat$x


check$lat <- for.longlat$y


check <- select(check,
                site, long, lat)

#merge with the sites data
sites <- merge(sites, all = T, check, by = "site")

#put data into the correct column and remove now extraneous columns

sites1 <- sites1 %>%
  mutate(
    longitude = ifelse(is.na(longitude) == TRUE, long, longitude),
    latitude = ifelse(is.na(latitude) == TRUE, lat, latitude)
  ) %>%
  select(-c(lat, long))

 
```


#rename variables
#filter species and records to exclude unknown species and moths, as well as surveys where no butterflies were seen

```{r}

#rename columns in species and...

species <- species %>% 
  rename( species.no = SPECIES, 
         common.name = COMMON_NAME,
         sci.name = SCI_NAME,
         strategy = STRATEGY,
         habitat = HABITAT) %>% 
  filter (is.na(include) == FALSE) #filter out non-butterflies and "unidentified" from species list

#so we can use it for summarising data etc later
species <- species %>% 
  mutate(species.no = as.character(species.no))

#rename columns in records
records<- records %>%
    rename(site = SITE, 
        species.no = SPECIES,
        year = YEAR,
        month = MONTH, 
        day = DAY, 
        week.no = WEEKNO, 
        day.no = DAYNO, 
        count = COUNT)%>% 
   mutate(species.no = as.character(species.no))

print("rows of records")
nrow(records)

records <- records %>% 
  filter (species.no %in% species$species.no) #this filters out records that were not butterflies or were not identified
      

print("rows of records")
nrow(records)


records <- records %>% 
  filter ( site %in% sites$site)


print("rows of records")
nrow(records)

```

```{r}

#create columns to identify survey events and site-year combinations
records<- records %>%
  mutate(survey.info = paste(site, year, month, day, sep = "_"), 
         site.year = paste(site, year, sep = "_"))

surveys <- records %>% 
  distinct(survey.info, .keep_all = TRUE)

#surveys per site per year
sum1 <-  surveys %>%
  group_by (site, year)  %>% 
summarise(n = n())
print("surveys per site per year")
summary(sum1$n)
hist(sum1$n, main = "Surveys per site per year")

sum2 <- sum1 %>%
  mutate(n = as.character(n)) %>%
  group_by(n) %>%
      summarise(freq.survey = n())
sum2

#get plot of sites surveyed per year by year
sites.per.year <- sum1 %>% 
  group_by(year)%>% 
summarise(sites = n())
print("sites per year")
summary(sites.per.year$sites)
plot(sites~year, data = sites.per.year)

#years per site summary and histogram
years.per.site <- sum1 %>% 
  group_by(site)%>% 
summarise(years = n())
print("years per site")
summary(years.per.site$years)
hist(years.per.site$years, xlim  = c(0, 50))



```
```{r}
site.and.year <- merge(sum1, sites, by = "site")

str(site.and.year$year)
```


```{r}


#x should be the above site.and.year
#condit is a condition to filter by


#filter_() might stop working, deprecated...
get.sites.in.date.range <- function(x, condit){
  sites.out <- x %>%
  filter_(condit)%>%
   select(site)%>%
      distinct

sites.out <- sites.out[['site']]
  return(sites.out)
}


sites.5.years <- get.sites.in.date.range(site.and.year, "year >= 2015 & year <=2019")


sites.1975 <- get.sites.in.date.range(site.and.year, "year == 1975")
sites.1980 <- get.sites.in.date.range(site.and.year, "year == 1980")
sites.1985 <- get.sites.in.date.range(site.and.year, "year == 1985")
sites.1990 <- get.sites.in.date.range(site.and.year, "year == 1990")
sites.1995 <- get.sites.in.date.range(site.and.year, "year == 1995")
sites.2000 <- get.sites.in.date.range(site.and.year, "year == 2000")
sites.2005 <- get.sites.in.date.range(site.and.year, "year == 2005")
sites.2010 <- get.sites.in.date.range(site.and.year, "year == 2010")
sites.2015 <- get.sites.in.date.range(site.and.year, "year == 2015")
sites.2016 <- get.sites.in.date.range(site.and.year, "year == 2016")
sites.2017 <- get.sites.in.date.range(site.and.year, "year == 2017")
sites.2018 <- get.sites.in.date.range(site.and.year, "year == 2018")
sites.2019 <- get.sites.in.date.range(site.and.year, "year == 2019")
sites.2022 <- get.sites.in.date.range(site.and.year, "year == 2022")


site.snapshots.to.map <- list(
  sites.1975,
  sites.1980, 
  sites.1985, 
  sites.1990, 
  sites.1995, 
  sites.2000, 
  sites.2005,
  sites.2010,
  sites.2015,
  sites.2018, 
  sites.2019, 
  sites.2022)


```

# records.site.year gives the number of records for each species in each site in each year

#records.site gives the total number of records for each species in each site over all years

```{r}



records.site.year <-  records %>%
  group_by (site, year, species.no)  %>% 
summarise(annual.records = sum(count))


#records.site.year.1 <- records.site.year

#records.site.year$site <- paste("site", records.site.year$site, sep = "_")
#records.site.year$species.no <- paste("species", records.site.year$species.no, sep = "_")


records.site <-  records %>%
  group_by (site, species.no)  %>% 
summarise(total.records = sum(count))
```
```{r}
head(records.site)
head(records.site.year)
```

# 2015-2019 data

```{r}

#to get the total records for each species in each site over  years:
records.site.5year <- records.site.year %>%
  filter(year %in% 2015:2019) %>% 
  group_by(site, species.no) %>%
  summarise(total.records = sum(annual.records))

records.site.5year$species.no <- paste("species", records.site.5year$species.no , sep = "_")

#get a site by species dataframe with counts of observations
site.by.obs.5.year.counts <- records.site.5year %>%
  pivot_wider(names_from = species.no,
              values_from = total.records) %>%
  column_to_rownames(var = "site")

# change NAs to zeros: these were cases 
site.by.obs.5.year.counts[is.na(site.by.obs.5.year.counts)] <- 0

#aggregate Thymelicus lineola and T.sylvestris records

"#due to difficulties in field identification, we aggregated Essex Skipper Thymelicus lineola and Small #Skipper Thymelicus sylvestris into one taxon) "

site.by.obs.5.year.counts <- site.by.obs.5.year.counts %>%
  mutate(species_1000 = species_119 + species_120) %>%
  select(-c(species_119, species_120))

#check if any sites have zero observations
any(rowSums(site.by.obs.5.year.counts) == TRUE)

site.by.obs.5.year.counts <- site.by.obs.5.year.counts[rowSums(site.by.obs.5.year.counts) > 0, ]

#make a presence absence site by species matrix
site.by.obs.5.year.presabs <- site.by.obs.5.year.counts

site.by.obs.5.year.presabs[site.by.obs.5.year.presabs[, ] > 0] <- 1
```

## make a row for combined species in species data 
```{r}

str(species)
#contents for new row
new.row <- c ("1000", "Combined Skippers", "Thymelicus lineola-sylvestris", "Wider countryside sp", "Grassland", 1)

#add new row and remove two old rows
species <- rbind(species, new.row) %>%
  filter(species.no != "119" & species.no !="120")
```



#check coverage
```{r}

#Transpose sity by counts matrix for use  by iNEXT package functions
check.coverage <- t(site.by.obs.5.year.counts)


#DataInfo() performs the first part of what iNEXT() does in seconds. the whole iNEXT() function would have to be left overnight for the whole dataset
#https://cran.r-project.org/web/packages/iNEXT/vignettes/Introduction.pdf

coverage <- DataInfo(check.coverage,  datatype="abundance")

hist(coverage$SC)

hist(coverage$SC, ylim = c(0, 50))


hist(coverage$SC, xlim = c(0.95, 1), breaks = 1000)

quantile(coverage$SC, seq(from = 0, to = 1, by = 0.01))


#it seems reasonable to include sites with greater than 0.95 coverage and surveyed over the flowering period

cov.merge <- coverage %>%
  select(Assemblage, n, S.obs, SC)

names(cov.merge) <- c("site", "total.butterflies", "richness", "coverage")

sites <- merge(sites, cov.merge, all.y = T)



```



#prepare additional site info from Rob
```{r}
sites.more.info <- sites.more.info %>%
  select(-c("Site Name", "Gridreference"))
 
names(sites.more.info) <- c("site", "easting", "northing", "length", "country", "no.sections", "no.yrs.surveyed", "first.year", "last.year", "survey.type" )



#interestingly, around 100 sites were not in 
sites <- merge(sites, sites.more.info, by = "site", all.x = TRUE)

sum3 <- sites %>%
  group_by(survey.type) %>%
  summarise(n = n(),
            richness = mean(richness),
            butterflies = mean(total.butterflies),
            coverage = mean(coverage)
      )
sum3
  
```

# obtain number of surveys
```{r}
sum1<- sum1 %>%
  filter(site %in% sites$site, 
         year %in% 2015:2019)

str(sum1)

#sites sampled per year
sum4 <- sum1 %>% 
  mutate(site = as.character(site),
         year = as.character(year))%>%
  group_by(site) %>%
  summarise(years.sampled = n(),
            total.surveys = sum(n),
            min.surveys = min(n))


sites <- merge(sites, sum4, by = "site")
```

```{r}
plot(richness~coverage, data = sites)

plot(coverage~total.surveys.x, data = sites)

plot(coverage~years.sampled.x, data = sites)

check <- filter(sites,
                coverage > 0.95,
                total.surveys.x >= 30,
                years.sampled.x >= 3)
```



#no longer needed

```{r}
#source(here("./Scripts/R_Scripts/get_site_by_species.R"))
```

#get site by species matrices for 2018
```{r}

# parallel::detectCores()

#get site by species matrix for 2018

#I am doing this for 2018 data only for now

# all.sites.in.a.period.data <- records.site.year %>%
#   ungroup() %>%  #can't perform the following operations on grouped data
#    filter(annual.records != 0, #my function assumes no row present when species is absent
#           year ==2018) %>% #choose a year
#             mutate(site = paste("site", site, sep = "_"),
#                      species.no = paste("species", species.no, sep = "_"))%>%
#           select(-year) 
# 
# 
# 
# site.by.species.2018 <- get.site.by.species(data = all.sites.in.a.period.data, sites = site, species = species.no, output = "presence.absence")


```


#get site by species matrices for 2015-2019
```{r}

# parallel::detectCores()

#get site by species matrix for 2015-2019

# all.sites.in.a.period.data <- records.site.year %>%
#   ungroup() %>%  #can't perform the following operations on grouped data
#    filter(annual.records != 0, #my function assumes no row present when species is absent
#           year >= 2015 & year <= 2019) %>% #choose a year
#               mutate(site = paste("site", site, sep = "_"),
#                      species.no = paste("species", species.no, sep = "_"), 
#                      year = as.character(species.no)) %>% 
#             group_by(site, species.no) %>%
#                 summarise(total.records = sum(annual.records))%>%
#   ungroup()
# 
# 
# site.by.species.2015.to.2019 <- get.site.by.species(data = all.sites.in.a.period.data, sites = site, species = species.no, output = "presence.absence")

#get the sites in all 5 years



```

#get the 
```{r}
#https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.12613


site.by.species.2015.to.2019.obs <- get.site.by.species(data = all.sites.in.a.period.data, sites = site, species = species.no, output = "obs.count")

site.by.species.2015.to.2019.obs <- site.by.species.2015.to.2019

```



#here I restrict the sites to only include those in all years. but actually, I intended to check coverage. working on this above
```{r}
all.year.sites <- sites.2015[sites.2015 %in% sites.2016]
all.year.sites <- all.year.sites[all.year.sites %in% sites.2017]
all.year.sites <- all.year.sites[all.year.sites %in% sites.2018]
all.year.sites <- all.year.sites[all.year.sites %in% sites.2019]

all.year.sites <- paste( paste("site", all.year.sites, sep = "_"))

site.by.species.2015.to.2019 <- filter(site.by.species.2015.to.2019, 
                                       rownames(site.by.species.2015.to.2019) %in% all.year.sites)
```




```{r}
#get year by species matrix for all sites with greater than 40 years of records

#prepare records data 
records.site.year.no.zeros <- records.site.year %>%
  ungroup() %>%  #can't perform the following operations on grouped data
   filter(annual.records != 0)  %>% #my function assumes no row present when species is absent
        mutate (year = as.character(year))
#get a vector of the chosen sites
use.sites <- years.per.site %>%
  filter(years >= 40)  %>%
    select(-years) %>%
     unlist() %>%
        as.numeric

#make a list to year by species matrices (which are actually dataframes)
year.by.species.matrices <- list()

# for loop to do it for each of the chosen sites and insert into list
  for (i in 1:length(use.sites)){
   one.site.over.years.data <- records.site.year.no.zeros %>%
      filter(site == use.sites[i]) 
     
   
one.site.over.years.data$species.no <- paste("species", one.site.over.years.data$species.no, sep = "_")
   
    year.by.species.matrices[[i]] <- get.site.by.species(data = one.site.over.years.data, sites = year, species = species.no, output = "presence.absence")
  }

#name the list elements after the sites
namevec <- vector()

length(namevec)<- length(use.sites)

  for (i in 1:length(use.sites)){
namevec[i] <- paste("site",  use.sites[i], sep = "_")
  }

names(year.by.species.matrices) <- namevec

#inspect output
str(year.by.species.matrices, max.level = 1)

check1 <- year.by.species.matrices[[1]]

head(year.by.species.matrices[[1]])

#below from package spaa not workign
##spaacheck <- data.frame(data2mat(check))
#data(testdata)
#spdf<- data.frame(data2mat(testdata))

#this would give a site by species matrix for all years combined
#site.by.species.all.years <- get.site.by.species(data = records.site, sites = site, species = species.no, output = "presence.absence")

#head(site.by.species.all.years)


```


```{r}
records.for.time <- records.site.year.no.zeros %>%
  filter(site %in% use.sites)%>%
  mutate(site.year = paste(site, year, sep = "_"))

check <- filter(records.for.time, year == 2000)

print(Sys.time())

 check.matrix <- get.site.by.species(data = records.for.time, sites = site.year, species = species.no, output = "presence.absence")

 print(Sys.time())
 
 
site.year.by.species <- get.site.by.species(data = records.for.time, sites = site.year, species = species.no, output = "presence.absence")
```


```{r}
surveys.site.year <- records%>%
  select (c(site, year, survey.info))  %>% 
    distinct()


surveys.site.year <- surveys  %>% 
  group_by(site, year)   %>% 
      summarise(surveys = n())
```


```{r}
records.40.years <- records %>%
  filter(site %in% use.sites, 
         count != 0)


check <- filter(records.40.years, year == 2010)

print(Sys.time())


```

```{r}
#stopCluster(cl)
```


# save processed data
```{r}
#write_rds(records.site, file = here("./Data/Processed_Data/records_site_year_all_years.rds"))


write_rds(site.year.by.species, file = here("./Data/Processed_Data/site_year_by_species.rds"))
write_rds(records, file = here("./Data/Processed_Data/processed_records.rds"))
write_rds(sites, file = here("./Data/Processed_Data/processed_sites.rds"))
write_rds(species, file = here("./Data/Processed_Data/processed_species.rds"))
write_rds(site.snapshots.to.map, file = here("./Data/Processed_Data/site_snapshots_to_map.rds"))
write_rds(use.sites, file = here("./Data/Processed_Data/sites_with_at_least_40_years.rds"))
write_rds(site.by.species.2015.to.2019, file = here("./Data/Processed_Data/site_by_species_2015_to_2019.rds"))

write_rds(site.by.species.2018, file = here("./Data/Processed_Data/site_by_species_2018.rds"))
write_rds(year.by.species.matrices, file = here("./Data/Processed_Data/year_by_species_matrices.rds"))
write_rds(surveys.site.year, file = here("./Data/Processed_Data/surveys_site_year.rds"))
```



```{r}


starwars %>% count(species)


starwars %>% 
  group_by(species) %>% 
    summarise(n = n())

head(starwars)


starwars %>% 
  group_by(species) %>% 
    summarise(totmass = sum(mass))
```




