---
title: "R Notebook"
output: html_notebook
---


#### Load packages.
```{r}
here::i_am("Scripts//Rmd_Scripts//temporal_zeta.Rmd")


library(tidyverse)
library(sf)
library(tmap)
library(zetadiv)
library(vegan)
library(gdm)
library(mgcv)
library(lme4)
library(here)
library(conflicted)
library(knitr)
library(knitrProgressBar)
library(tictoc)

conflicts_prefer(dplyr::select)
conflicts_prefer(knitrProgressBar::progress_estimated)

conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::mutate)
conflicts_prefer(dplyr::select)
conflicts_prefer(base::split)
```


#### Functions
```{r}
source(here("Scripts/R_scripts/butterfly_richness_turnover_functions.R"))
```

#### File Paths
```{r}
# Path to data used for tje msgdm analysis (which is already done)
data_path <- paste0("./Data/Processed_Data/Temporal/temporal_ukbms_obs_sites_spp.rds")


# Path to folder to save output of this script
out_path <- here(paste0("./Output/Temporal/Retreat/"))

if(!dir.exists(out_path)) dir.create(out_path, recursive = TRUE)

# Basemap without outlying islands
GB_simpl <- readRDS(here("./Data./Maps./gb_multipolygon_simplified.rds"))
```

#### Data
```{r}
data <- readRDS(here(data_path))

site.by.species <- data$site.year.by.species.presabs


site.by.species.counts <- data$site.year.by.species.counts
  
#site.by.env <- data$site.by.env

species <- data$species

sites <- data$sites 

site.by.xy <- sites %>%
  select(site, x, y)  %>%
  mutate(site = paste0("site_", site)) %>%
  column_to_rownames(var = "site")
```


# to do, include the matrix needed for plots below in data

#### Inspect coverage for time series
```{r}
# site.year.cov <- site.year.cov[order(site.year.cov$year, decreasing = FALSE), ] 
# 
# cov_mat <- site.year.cov %>%
#   select(site, year, SC) %>%
#   pivot_wider(values_from = SC, names_from = site) %>%
#   column_to_rownames(var = "year") %>%
#   as.matrix()
# 
# dim(cov_mat)
# 
# heatmap(cov_mat[, 1:50], Rowv= NA, Colv = NA)
# heatmap(cov_mat[, 51:100], Rowv= NA, Colv = NA)
# heatmap(cov_mat[, 101:150], Rowv= NA, Colv = NA)
# heatmap(cov_mat[, 151:201], Rowv= NA, Colv = NA)
# heatmap(cov_mat[, 151:ncol(cov_mat)], Rowv= NA, Colv = NA)
# # I am not sure why, but 1973-1975 data appear empty in plot, but they are there in the input

```



#### Plot Sites
```{r}
plot.points <- st_as_sf(sites, coords=c("x","y"), remove = FALSE, crs = 27700)

basemap <-tm_shape(GB_simpl) +
  tm_borders()

pointsmap <- basemap +
tm_shape(plot.points) +
  tm_symbols(size = 0.05, col = "blue")

pointsmap2 <- basemap +
tm_shape(plot.points) +
  tm_symbols(size = 0.5, alpha = 0.5, col = "blue")

print(pointsmap)
print(pointsmap2)

filepath <- paste0(out_path, "/temporal_sites_map.png")

png(file = filepath)
print(pointsmap2)
dev.off()
```


# Get richnes for each site in each year

```{r}
rich.data <- data.frame(
  site = str_split_i(rownames(site.by.species), "_", 2),
  year = str_split_i(rownames(site.by.species), "_", 3),
  richness = rowSums(site.by.species)
)

rich.data$year <- as.numeric(rich.data$year)
str(rich.data)
```
# Get years per site, start and end year of each site, and sites per year
```{r}
years.per.site <- rich.data %>%
  group_by(site) %>%
  summarize(years = n(),
            start.year = min(year),
            end.year = max(year),
            span = end.year - start.year)

sites.per.year <- rich.data %>%
  group_by(year) %>%
  summarize(sites = n())

# years for sites
ggplot(years.per.site, aes(x = years)) + 
  geom_histogram()

filepath <- paste0(out_path, "/temporal_sites_start_year.png")
png(file = filepath)
ggplot(years.per.site, aes(x = start.year)) + 
  geom_histogram()
dev.off()

# span for sites
ggplot(years.per.site, aes(x = span)) + 
  geom_histogram()

filepath <- paste0(out_path, "/temporal_sites_span.png")
png(file = filepath)
ggplot(years.per.site, aes(x = start.year)) + 
  geom_histogram()
dev.off()

# Start year for sites
ggplot(years.per.site, aes(x = start.year)) + 
  geom_histogram()

filepath <- paste0(out_path, "/temporal_sites_end_year.png")
png(file = filepath)
ggplot(years.per.site, aes(x = end.year)) + 
  geom_histogram()
dev.off()

# End year for sites
ggplot(years.per.site, aes(x = end.year)) + 
  geom_histogram()

filepath <- paste0(out_path, "/temporal_sites_years_per_site.png")
png(file = filepath)
ggplot(years.per.site, aes(x = years)) + 
  geom_histogram()
dev.off()

# Sites per year
ggplot(sites.per.year, aes(year, sites)) + 
  geom_point()


filepath <- paste0(out_path, "/temporal_sites_per_year.png")
png(file = filepath)
ggplot(sites.per.year, aes(year, sites)) + 
  geom_point()
dev.off()

```

# Examples of richness time series
```{r}
rich.data$site.num <- as.numeric(rich.data$site)

d2 <- filter(rich.data, site.num <= 10)

ggplot(d2, aes(year, richness)) + 
  geom_point() + 
  facet_wrap(~ site) + 
  geom_smooth()

filepath <- paste0(out_path, "/temporal_sites_eight_examples_richness.png")
png(file = filepath)
ggplot(d2, aes(year, richness)) + 
  geom_point() + 
  facet_wrap(~ site) + 
  geom_smooth()
dev.off()

```
# Get pairwise similarity or dissimilarity via gdm::formatsitepair

```{r}
data(varespec)
vare.dist <- vegdist(varespec)

head(varespec)

vare.dist[1:10]

data(sipoo)
m <- betadiver(sipoo)
plot(m)
## The indices
betadiver(help=TRUE)
## The basic Whittaker index
d <- betadiver(sipoo, "w")


x <- matrix(rnorm(100), nrow = 5)
dist(x)
dist(x, diag = TRUE)
dist(x, upper = TRUE)
m <- as.matrix(dist(x))

v <- x %>%
  dist() %>%
  as.matrix() %>%
  data.frame()

(m2 <- matrix(1:20, 4, 5))
m3 <- lower.tri(m2)
 m2 *m3

```
```{r}
#-----------------------------------------------
# 1) Extract the lower triangular part (LTP)
#    of matrix as a vector
#-----------------------------------------------
(mat <- matrix(1:25, nrow = 5))
 
# column-wise
(v_cw <- mat[lower.tri(mat, diag = FALSE)] )
# row-wise
(v_rw <- c(t(mat))[c(t(lower.tri(mat, diag = FALSE)))] )
```


```{r}


 
cat("Input matrix:\n"); print(mat)
 
# vector containing the LTP
print(v_cw)
print(v_rw)
 
#-----------------------------------------------
# 2) Insert a vector into 
#    the lower triangular part of matrix
#-----------------------------------------------
A <- B <- matrix(0,3,3)
 
# column-wise
A[lower.tri(A, diag = TRUE)] <- v_cw
# row-wise
B[t(lower.tri(B, diag = TRUE))] <- v_rw; B <- t(B)
 
# reconstructed lower triangular matrix 
print(A) 
print(B) 

# I need to turn a triangle from distance matrix into a vector
# I need years corresponding to rows and columns for each distance

x1 <- 1:10
M <- matrix(0,5,5)
M[lower.tri(M, diag=FALSE)] <- x1


t(combn(1:5, 2))
```


# format environmental data for gdm package
```{r}
split_factor <- str_split_i(rownames(site.by.species), "_", 2)

set.seed(1)
site.by.env <- data.frame(
rand1 = runif(n = nrow(site.by.species)),
rand2 = rnorm(n = nrow(site.by.species))
)

head(site.by.env)


```

# Format site by species count data for gdm
```{r}
site.by.species.counts <- site.by.species.counts %>%
  mutate(x = as.numeric(str_split_i(rownames(site.by.species.counts), "_", 3)),
         y = 0,
         .before = species_100)
  
head(site.by.species.counts)    


# for gdm
# site.by.add.in <- cbind(site.by.env, site.by.xy) %>%
#   mutate(site = str_split_i(rownames(site.by.add.in), "_", 2)) %>%
#   select(site, year, x, y)
#   
# site.by.species.counts <- cbind(site.by.add.in, site.by.species.counts)

```

# Format site by species count data for msgdm
```{r}
site.by.species.presabs <- data$site.year.by.species.presabs %>%
  mutate(site.year = rownames(site.by.species.counts), .before = species_100)
 

head(site.by.species.counts)
head(site.by.env)

```


# Split data
```{r}

# dataset will be split by year anyway
env.list<- site.by.env %>% 
  split(split_factor)

# get data for gdm as lists just use example ones
 counts.list <- split(
  site.by.species.counts, split_factor)
 


#counts.list <- counts.list[c("1", "10", "2", "3", "4", "6", "8", "9")]
# env.list <- env.list[c("1", "10", "2", "3", "4", "6", "8", "9")]
#rand_samp <- sample(1:nrow(site.by.env), 100, replace = FALSE)

```


#still need to check simpson distance calculated and appended as intended (right values, right order)

# Note, estimated abundance indices are available and are supposed to be more accurate
# here we get Bray Curtis

# idea is to add other indices and have them available in same dataframe
```{r}

pairwise_list <- list()

for (i in 1:length(counts.list)){  

  site.by.env.i <- env.list[[1]]%>%
    mutate(year.id = 1:nrow(.))

  site.by.species.counts.i <- counts.list[[1]] %>%
    mutate(year.id = 1:nrow(.), .before = species_100)

# use formatsitepair() to get bray-curtis distanc
  formatted.data.1 <- formatsitepair(bioData = site.by.species.counts.i, bioFormat = 1, dist = "bray", abundance = TRUE, 
                       siteColumn = "year.id", XColumn = "x", YColumn = "y", predData = site.by.env.i)

# use formatsitepair to get horn-morisita distance
  formatted.data.2 <- formatsitepair(bioData = site.by.species.counts.i, bioFormat = 1, dist = "horn", abundance = TRUE, 
                       siteColumn = "year.id", XColumn = "x", YColumn = "y", predData = site.by.env.i)

# get simpson distance   
 d <- designdist(site.by.species.counts.i, method = "1-(J)/(J = min(A+B))",
         terms = c("binary"), 
         abcd = FALSE, alphagamma = FALSE, "simpson_distance", maxdist = 1)
   
   # add other distances to the first data frame
   formatted.data.1 <- formatted.data.1 %>%
    rename(bc.distance = distance) %>%
    mutate(
     bc.similarity = 1 - bc.distance,
     hm.distance = formatted.data.2$distance,
     hm.similarity = 1 - hm.distance,
     simpson.distance = d,
     simpson.similarity = 1 - d,
     time.distance = abs(s1.xCoord - s2.xCoord),
     .before = weights)

 pairwise_list[[i]] <- formatted.data.1
 
}

names(pairwise_list) <- unique(split_factor) #used as entries for site below

pairwise_distances <- bind_rows(pairwise_list, .id = "site")

head(pairwise_distances)

```

```{r}
plot(bc.similarity ~ hm.similarity, data = pairwise_distances)
plot(bc.similarity ~ simpson.similarity, data = pairwise_distances)
```


# check what is still needed below


# get pairwise diversity
```{r}

site.species.list <- split(
  site.by.species, split_factor)

zeta.time <- data.frame(
  site = factor(levels = unique(split_factor)),
  zeta = numeric(), 
  time.diff = numeric())


# for (i in 1:length(site.species.list)){
#   year.by.species <- site.species.list[[i]]
#   year.xy <- data.frame(x = 1:nrow(year.by.species), y = 0)
#   site <- str_split_i(rownames(year.by.species[1,]), "_", 2)
#   
#   zeta.ddecay.sample <- Zeta.ddecay(
#     xy = year.xy, 
#     data.spec = year.by.species, 
#     order = 2, 
#     #normalize = "Simpson",
#     reg.type="glm", 
#     sam = 5000)
#   out.data <- data.frame(
#     site,
#     zeta.ddecay.sample$zeta.val, 
#     zeta.ddecay.sample$distance)
#  zeta.time <- rbind(zeta.time, out.data)
#   print(i)
# }

names(zeta.time) <- c("site", "zeta.val", "time.diff")
zeta.time$site <- as.factor(zeta.time$site )


head(site.by.species.counts)

head(rich.data)


```



```{r}


filepath <- paste0(out_path, "/examples_temporal_turnover.png")

# to have the same sites as for richness example
site.species.list.eight <- site.species.list[c("1", "10", "2", "3", "4", "6", "8", "9")]

png(file = filepath)
par(mfrow = c(3,3))
for (i in 1:8){
  year.by.species <- site.species.list.eight[[i]]
  year.xy <- data.frame(x = 1:nrow(year.by.species), y = 0)
  site <- str_split_i(rownames(year.by.species[1,]), "_", 2)
  
  zeta.ddecay.sample <- Zeta.ddecay(
    xy = year.xy, 
    data.spec = year.by.species, 
    order = 2, 
    #normalize = "Simpson",
    reg.type="gam", 
    sam = 5000)
  out.data <- data.frame(
    site,
    zeta.ddecay.sample$zeta.val, 
    zeta.ddecay.sample$distance)
 zeta.time <- rbind(zeta.time, out.data)
  print(i)
}



filepath <- paste0(out_path, "/examples_temporal_turnover_simpson.png")

# to have the same sites as for richness example
site.species.list.eight <- site.species.list[c("1", "10", "2", "3", "4", "6", "8", "9")]

png(file = filepath)
par(mfrow = c(3,3))
for (i in 1:8){
  year.by.species <- site.species.list.eight[[i]]
  year.xy <- data.frame(x = 1:nrow(year.by.species), y = 0)
  site <- str_split_i(rownames(year.by.species[1,]), "_", 2)
  
  zeta.ddecay.sample <- Zeta.ddecay(
    xy = year.xy, 
    data.spec = year.by.species, 
    order = 2, 
    normalize = "Simpson",
    reg.type="gam", 
    sam = 5000)
  out.data <- data.frame(
    site,
    zeta.ddecay.sample$zeta.val, 
    zeta.ddecay.sample$distance)
 zeta.time <- rbind(zeta.time, out.data)
  print(i)
}
dev.off()


```


```{r}
gls.rich <-gls(richness ~ year, method="REML",
             correlation = corCompSymm(form= ~ 1 | site),
            data = rich.data)
summary(gls.rich)

anova(gls.rich)
```


```{r}
g2 <- ggplot(rich.data,aes(x = year,y = richness, colour = site))+ 
		theme_classic() 

# Plotting multiple Regression Lines 
g2 + geom_line(
  stat = "smooth",
  method = "lm",
  se = FALSE,
  fullrange = FALSE, 
	aes(
	  color = site, 
	  alpha = 0.5))  +
  
  theme(legend.position = "none")
```

```{r}
gls.fit<-gls(zeta.val ~ time.diff, method="REML",
             correlation = corCompSymm(form= ~ 1 | site),
            data = zeta.time)
summary(gls.fit)

anova(gls.fit)
```





```{r}


g1 <- ggplot(zeta.time,aes(x = time.diff,y = zeta.val, colour = site))+ 
		theme_classic() 

# Plotting multiple Regression Lines 
g1 + geom_line(
  stat = "smooth",
  method = "lm",
  se = FALSE,
  fullrange = FALSE, 
	aes(
	  color = site, 
	  alpha = 0.5))  +
  
  theme(legend.position = "none")

```

```{r}
m1 <-lm(zeta.val ~ time.diff, data =  zeta.time)
summary(m1)
```


```{r}
m2 <-lm(zeta.val ~ site*time.diff, data =  zeta.time)

anova(m2)
#summary(m2)
```



















## Zeta Declines

```{r}
head(site.by.species)
```



#### Get a site
```{r}
 d1 <- site.by.species %>%
   rownames_to_column(var = "site.year") %>%
   filter(grepl('site_16_', site.year))  %>% #filter to site 1
   column_to_rownames(var = "site.year")

 years <- str_split_i(rownames(d1), "_", 3) %>%
   as.integer()
```

#### Orders 1:10
```{r}


zd <- Zeta.decline.ex(data.spec = d1, orders = 1:30)
# "Power Law"
summary(zd$zeta.pl)
# Exponential")
summary(zd$zeta.exp)
# "AIC"
# zd$aic has dollar signs in rownames, which messes up resulting table from kable
# therefore I edit the rownames
zd_performance <- zd$aic
rownames(zd_performance) <- c("Exp_to_zeta_10", "PL_to_zeta_10")
zd_performance$R.sq.multi <- c(summary(zd$zeta.exp)$r.squared, 
                              summary(zd$zeta.pl)$r.squared)
zd_performance$R.sq.adj <- c(summary(zd$zeta.exp)$adj.r.squared, 
                              summary(zd$zeta.pl)$adj.r.squared)
kable(zd_performance)

# Keep to add values from below
all_performance <- zd_performance
```

```{r}
year.xy <- data.frame(x = 1:50, y = 0)
```


# this one uses first year and zeta diversity is species in common between the first i years
```{r}


zd <- Zeta.decline.mc(data.spec = d1, xy = year.xy, FPO = c(1, 0), DIR = FALSE, orders = 1:30, sam = 10)#sam not used

```

# This one uses a directed nearest neighbour approach, site.years are sampled in NN sets but not always starting at first year

```{r}


year.xy <- year.xy[1:nrow(d1), ]

set.seed(1)
zd <- Zeta.decline.mc(data.spec = d1, xy = year.xy, FPO = c(1, 0), DIR = TRUE, orders = 1:30, sam = 1000)

set.seed(1)
zd <- Zeta.decline.mc(data.spec = d1, xy = year.xy, FPO = c(1, 0), DIR = TRUE, orders = 1:10, sam = 1000)
```



```{r}
set.seed(1) 

zeta.ddecays <- Zeta.ddecays(xy = year.xy, data.spec = d1, sam = 1000, orders = 2:9,
plot = TRUE, confint.level = 0.95)
```

```{r, warning = FALSE, label='Zeta Distance Decay GAMs'}
set.seed(1) 
zeta.ddecay.gam <- Zeta.ddecay(xy = year.xy, data.spec = d1, order = 2, 
                                    reg.type="glm", sam = 5000)

set.seed(1) 
zeta.ddecay.gam <- Zeta.ddecay(xy = year.xy, data.spec = d1, order = 3, 
                                    reg.type="glm", sam = 5000)
length(zeta.ddecay.gam$distance)

check <- data.frame(cbind(zeta.ddecay.gam$zeta.val, zeta.ddecay.gam$distance))
set.seed(1) 
zeta.ddecay.gam <- Zeta.ddecay(xy = year.xy, data.spec = d1, order = 5, 
                                    reg.type="glm", sam = 5000)

set.seed(1) 
zeta.ddecay.gam <- Zeta.ddecay(xy = year.xy, data.spec = d1, order = 8, 
                                    reg.type="glm", sam = 5000)

set.seed(1) 
zeta.ddecay.gam <- Zeta.ddecay(xy = year.xy, data.spec = d1, order = 10, 
                                    reg.type="glm", sam = 5000)


```

```{r}

data$site.by.env <- site.by.env

# Write for Cang

cang_path <- here("./Data/Processed_Data/Temporal/Cang/")


site.by.xy <- site.by.xy %>%
  mutate(site = rownames(site.by.xy), .before = x)

site.by.xy <- merge(site.by.env, site.by.xy, by = "site")

site.by.xy <- site.by.xy %>% 
  mutate(site.year = paste0(site, "_" ,year), .before = site)

site.by.xy <- site.by.xy %>%
  select(site.year, site, year, x, y)

write_csv(site.by.species.counts, file = paste0(cang_path, "/",   "site_by.species.counts.csv"))

write_csv(site.by.species.presabs, file = paste0(cang_path, "/",   "site_by.species.presabs.csv"))


write_csv(site.by.xy, file = paste0(cang_path, "/",   "site_by.xy.csv"))
head(site.by.xy)

site.by.env <- site.by.xy %>%
  column_to_rownames(var = "site.year")%>%
  select(year)

site.by.xy <- site.by.xy %>%
  column_to_rownames(var = "site.year")%>%
  select(x, y)



site.by.species.presabs <- as.integer(site.by.species.presabs)

c1 <- data.frame(sapply(site.by.species.presabs, FUN = as.integer))

c2 <- c1[1:500,]

```


```{r}
set.seed(1)
zm1 <- Zeta.msgdm(xy = site.by.xy, data.spec = c1, data.env = site.by.env,  order = 2, sam = 50000,
  reg.type="ispline", normalize = "Simpson", family=binomial(link="log"),cons.inter = -1, glm.init = TRUE) 

Plot.ispline(msgdm = zm1, data.env = site.by.env, distance = TRUE,  legend = TRUE)

summary(zm1$model)

```

```{r}

site.by.env2 <- select(site.by.env, year)
set.seed(1)
zm2 <- Zeta.msgdm(xy = site.by.xy, data.spec = c1, data.env = site.by.env2,  order = 2, sam = 50000,
  reg.type="ispline", normalize = "Simpson", family=binomial(link="log"),cons.inter = -1, glm.init = TRUE) 

Plot.ispline(msgdm = zm2, data.env = site.by.env2, distance = TRUE,  legend = TRUE)

summary(zm2$model)

```



```{r}

set.seed(1)
zm3 <- Zeta.msgdm(xy = site.by.xy, data.spec = c1, data.env = site.by.env2,  order = 2, sam = 50000,
  reg.type="ispline",  cons.inter = -1, glm.init = TRUE) 

Plot.ispline(msgdm = zm3, data.env = site.by.env2, distance = TRUE,  legend = TRUE)

summary(zm3$model)

```
# , use gdm package to get pairwise dissimilarity

```{r}


ggplot(bc1, aes(time.distance, BC_simillarity)) + 
  geom_point() +
  facet_wrap(~ site) + 
  geom_smooth()


filepath <- paste0(out_path, "/temporal_sites_eight_examples_bc.png")
png(file = filepath)
ggplot(bc1, aes(time.distance, BC_simillarity)) + 
  geom_point() +
  facet_wrap(~ site) + 
  geom_smooth()
dev.off()


```
# attempt at gdm, check predictors

```{r}
gdm.1 <- gdm(data = pairwise_list[[1]][, 1:12], geo=TRUE)
 
 length(gdm.1$predictors) # get ideal of number of panels

summary(gdm.1)
plot(gdm.1, plot.layout=c(2,3))

```


#gdm example
```{r}
# start with the southwest data table

 sppData <- southwest[c(1,2,13,14)]
 envTab <- southwest[c(2:ncol(southwest))]
 head(envTab)
 head(sppData)
 
#########table type 1
## site-species table without coordinates
testData1a <- reshape2::dcast(sppData, site~species)
head(testData1a)
##site-species table with coordinates
coords <- unique(sppData[, 2:ncol(sppData)])
testData1b <- merge(testData1a, coords, by="site")
## site-species, table-table
exFormat1a <- formatsitepair(testData1a, 1, siteColumn="site", XColumn="Long",
YColumn="Lat", predData=envTab)
head(testData1a)
head(envTab)
```
